---
title: "Chytrid_meta_analysis"
---
```{r}
install.packages(c(
  "dplyr",
  "tidyr",
  "readr",
  "pacman",
  "metafor",
  "corrplot",
  "ape",
  "stringr",
  "minqa"
))

library(dplyr)
library(tidyr)
library(readr)
library(metafor)
library(corrplot)
library(ape)
library(stringr)
library(minqa)

install.packages("remotes")
remotes::install_github("daniel1noble/orchaRd")
library(orchaRd)
```

### Prevalence (log risk ratio) ###
### Main analysis ###
```{r}
## read.table
metadata <- read.csv("metadata.csv")
preval <- metadata %>% filter(trait == "prevalence")

## create new columns
preval$T_all <- round(preval$n_T_animals)
preval$T_infected <- round(preval$mean_T_trait/100 * preval$n_T_animals)
preval$T_uninfected <- round(preval$n_T_animals - preval$T_infected)

preval$C_all <- round(preval$n_C_animals)
preval$C_infected <- round(preval$mean_C_trait/100 * preval$n_C_animals)
preval$C_uninfected <- round(preval$n_C_animals - preval$C_infected)

preval <- preval %>% filter(!(T_infected == 0 & C_infected == 0)) ## remove column when all groups have prevelence = 0 %           
#############################
## log risk ratio (log RR) ##
#############################

preval_es <- escalc(
  measure = "RR",
  ai = T_infected,  bi = T_uninfected, # ai = event, bi = non-event
  ci = C_infected,  di = C_uninfected, # ci = event, di = non-event
  data = preval,
  add = 0.5, to = "only0") ## add 0.5 when a cell is zero 

# This prevents the log RR from being infinite (when % in treatment = 0) or undefined (when % in control = 0)) 

###########################################
## Non-independence: Shared animal group ##
###########################################

# Adjust non-independence of sampling variances/co-variances
V_preval <- vcalc(vi      = vi,
           cluster = study_exp_id, # experiment level where shared value occurred
           grp1    = T_ID, # cohort ID of treatment
           grp2    = C_ID, # cohort ID of control
           w1      = n_T_animals,
           w2      = n_C_animals,
           data = preval_es)

# Visualize
corrplot::corrplot(cov2cor(V_preval)) 

######################################
## Random factor: observation-level ##
######################################

preval_es$obs <- 1:nrow(preval_es)

##############################
## Random factor: phylogeny ##
##############################

tree <- read.tree("tree_for_imputation.tre")

# Species set from my metadata
sp_data <- preval_es$host_species %>% as.character() %>%
  str_squish() %>%           # collapse multiple spaces
  str_trim()

sp_data_treefmt <- gsub(" ", "_", sp_data)  # match typical tree tip style

# Check any mismatch between species in the metadata and species in existing tree
sp_tree_spaces <- gsub("_", " ", tree$tip.label)
setdiff(unique(sp_data), sp_tree_spaces) # In data but not in tree

# Prune the tree: only keep species present in BOTH tree and my metadata
tips_to_keep <- intersect(tree$tip.label, unique(sp_data_treefmt))
tree_pruned_preval <- keep.tip(tree, tips_to_keep) 

# Force the tree to be binary
tree_pruned_preval <- multi2di(tree_pruned_preval)
is.binary (tree_pruned_preval) # yes, the tree is binary

# visualize the tree
plot(tree_pruned_preval, type = "radial", cex = 0.8, no.margin = TRUE)

# Create correlation phylogeny matrix for analysis
phylo_cor_preval <- vcv(tree_pruned_preval, corr = TRUE)

# Create a column in metadata that matches the tree’s tip-label format
preval_es <- preval_es %>%
  mutate(species_OTL = gsub(" ", "_", str_squish(as.character(host_species))),
         species_OTL = factor(species_OTL))

####################
## Overall effect ##
####################

M3_preval <- rma.mv(yi = yi, V = V_preval,
  mods = ~  1,
  random = list(~1 | study_id, ~1 | species_OTL, ~1 | obs), 
  R = list(species_OTL = phylo_cor_preval),
  data = preval_es,
  method = "REML", 
  test = "t") 

summary(M3_preval)

## Estimate absolute heterogenesity
sum(M3_preval$sigma2)
## Estimate relative heterogeneity
orchaRd::i2_ml(M3_preval)
orchaRd::cvh2_ml(M3_preval)
orchaRd::m2_ml(M3_preval)


# Count the number of studies, experiments and species
preval_es %>%
  summarise(n_k = n(),
            n_study = n_distinct(study_id),
            n_species = n_distinct (species_OTL),
            n_exp = n_distinct (study_exp_id),
            .groups = "drop")  

##########################
## Procedural variation ## 
##########################

#### ** (1) Measurement effort throughout the experiment ##
# Assuming a much lower prevalence in treated group than in controls with increasing measurement effort 
# One more measure → negative log RR (treatment vs control).

# standardize (mean-centered) n_infection_measure
preval_es$n_infection_measure_c <- scale(preval_es$n_infection_measure, center = TRUE, scale = FALSE)

#### ** (2) Time elapsed between infection challenges and outcome measurement ##

# Assume higher prevalence risk in controls than in treated groups when prevalence is measured long after Bd exposure
# (i.e., more time for disease to develop, so more control individuals were infected).
# A one-day increase in elapsed time → negative log RR (treatment vs control).

# Time elapsed has been extracted separately for control and treatment groups. Two values are highly correlated (because they often measured together) so cannot be treated as independent moderators. I took the average time_elapsed.

# standardize (mean-centered) average time at measure 
preval_es$avg_time_measure   <- (preval_es$T_time_post + preval_es$C_time_post)/2
preval_es$time_measure_c <- scale(preval_es$avg_time_measure, center = TRUE, scale = FALSE)

#### ** (3) Time elapsed between immunisation and infection challenge ##

# Assuming increased risk of being infected in treated groups when pathogen re-exposure occurs long after completion of treatment 
# (i.e., treatment-induced protection wanes over time)
# A one-day increase in the time interval → positive log RR (treatment vs control).

# standardize (mean-centered) time intervals
preval_es$time_persist_c <- scale(preval_es$time_interval, center = TRUE, scale = FALSE)

# MODEL accounting for (1) measurement effort (2) time intervals between challenge and measurement and (3) time interval between immunisation and challenge
M3_cov_pre <- rma.mv(yi = yi, V = V_preval,
  mods = ~  1 + time_persist_c + time_measure_c + n_infection_measure_c,
  random = list(~1 | study_id, ~1 | species_OTL, ~1 | obs), 
  R = list(species_OTL = phylo_cor_preval),
  data = preval_es,
  method = "REML", 
  test = "t") 

summary(M3_cov_pre) ## none of these procedural variables were significant. I excluded them to test for the moderating effects of immunisation type and host type.

#############################################
## categorical variable: immunisation type ##
#############################################

# Set as a factor
preval_es$treatment <- factor(
  preval_es$treatment,
  levels = c("other","synthetic_antiparastics", "dead_pathogen", "probiotic","live_pathogen","natural_chem"))

## ANALYSIS: test for the effect for individual treatment

M_treat_infected <- rma.mv(yi = yi, V = V_preval,
  mods = ~  -1 + treatment,
  random = list(~1 | study_id, ~1 | species_OTL, ~1 | obs), 
  R = list(species_OTL = phylo_cor_preval), data = preval_es,
  method = "REML", test = "t") 

summary(M_treat_infected) 

## Prediction intervals
levs <- levels(preval_es$treatment)
out <- cbind(treatment = levs,
             round(as.data.frame(predict(M_treat_infected,
                                         newmods = model.matrix(~ -1 + treatment,
                                                               data = data.frame(treatment = levs)))), 3))
out

# Check k per treatment type
preval_es %>%
  group_by(treatment) %>%
  summarise(total_k = n(),
            .groups = "drop")  

######################################
## categorical variable: life stage ##
######################################

## There are two types of life stage:
## 'life_stage_treated' at which prophylatic treatments were applied (= during treatment)
## 'life_stage_tested' at which pathogen resistance was tested (= during pathogen exposure)

## I first assessed how many studies having mismatch between the two life stages (= cross-stage study)
preval_es %>%
  group_by(life_stage_matched) %>%
  summarise(total_n = n(),
            .groups = "drop") # two life stages are the same in most of the effect sizes (9/90 = 90%)

# Therefore, I only accounted for life_stage_tested
# Set as a factor
preval_es$life_stage_tested <- factor(
  preval_es$life_stage_tested,
  levels = c("adult","juvenile", "larva"))

# Check k per treatment type
preval_es %>%
  group_by(life_stage_tested) %>%
  summarise(total_k = n(),
            .groups = "drop")  
            
## ANALYSIS: test for the effect for individual life stage
M_stage_infected <- rma.mv(yi = yi, V = V_preval,
  mods = ~  -1 + life_stage_tested,
  random = list(~1 | study_id, ~1 | species_OTL, ~1 | obs), 
  R = list(species_OTL = phylo_cor_preval), data = preval_es,
  method = "REML", test = "t", control = list(optimizer = "bobyqa")) 

summary(M_stage_infected) 

## Prediction intervals
levs <- levels(preval_es$life_stage_tested)
out_stage <- cbind(life_stage_tested = levs,
                   as.data.frame(predict(M_stage_infected,
                     newmods = model.matrix(~ -1 + life_stage_tested,
                       data = data.frame(life_stage_tested = levs)))))
out_stage

## Estimate absolute heterogenesity
sum(M_stage_infected$sigma2)
## Estimate relative heterogeneity: I^2 (orchaRd)
orchaRd::i2_ml(M_stage_infected)

#######################################
## categorical variable: host origin ##
#######################################

# 'Larvae' were all sourced from the wild. I removed "larva" group here and test the effect of host origin (wild vs captive) for "adult" and "juvenile" only.

# Keep only adult + juvenile rows
keep_preval <- preval_es$life_stage_tested %in% c("adult","juvenile")

# Subset the already-built variance-covariance matrix
V_keep_preval <- V_preval[keep_preval, keep_preval]

# Subset the data (random effects & phylo R will align by factor levels automatically)
data_preval <- preval_es[keep_preval, ]

# Set as a factor
data_preval$host_origin <- factor(
  data_preval$host_origin,
  levels = c("captive", "wild"))

# ANALYSIS: test for the effect for individual host origin
M_host_infected <- rma.mv(yi = yi, V = V_keep_preval,
                         mods = ~ -1 + host_origin,
                         random = list(~1 | study_id, ~1 | species_OTL, ~1 | obs), 
                         R = list(species_OTL = phylo_cor_preval), data = data_preval,
                         method = "REML", test = "t") 

summary(M_host_infected)

# Count k per origin
data_preval %>%
  group_by(host_origin) %>%
  summarise(total_n = n(),
            .groups = "drop") 


#########################################
## categorical variable: host taxonomy ##
#########################################

# Set as a factor
preval_es$family <- factor(preval_es$family)

preval_es$family <- factor(
  preval_es$family,
  levels = c("Bufonidae", "Ranidae", "Hylidae", "Alytidae", "Cryptobranchidae", "Sirenidae", "Plethodontidae", "Pipidae", "Ambystomatidae"))

## ANALYSIS: test for the effect for individual family

M_taxa_infected <- rma.mv(yi = yi, V = V_preval,
  mods = ~  -1 + family,
  random = list(~1 | study_id, ~1 | species_OTL, ~1 | obs), 
  R = list(species_OTL = phylo_cor_preval), data = preval_es,
  method = "REML", test = "t") 

summary(M_taxa_infected) 

## Prediction intervals
levs <- gsub("^family", "", names(coef(M_taxa_infected)))  # nice labels (optional)

Xnew <- diag(length(coef(M_taxa_infected)))
colnames(Xnew) <- names(coef(M_taxa_infected))

pred_taxa <- predict(M_taxa_infected, newmods = Xnew)
options(digits = 3)
out_taxa <- data.frame(family = levs, as.data.frame(pred_taxa))
out_taxa

# Check k per family type
preval_es %>%
  group_by(family) %>%
  summarise(total_k = n(),
            .groups = "drop")  
          
```
## Prevalence ##
## checking for publication bias ##
```{r}
# Standardize (mean-centered) publication year
preval_es$year_c <- scale(preval_es$publication_year, center = TRUE, scale = FALSE)

# Calculate the inverse of the "effective sample size" 
preval_es$inv_n_tilda <-  with(preval_es, (n_T_animals + n_C_animals)/(n_T_animals*n_C_animals))
preval_es$sqrt_inv_n_tilda <-  with(preval_es, sqrt(inv_n_tilda))

# sqrt (inverse of effective sample size) = uncertainty / small sample size => larger value = smaller N, more uncertain

# Add sqrt__inverse_effective_sample_size to the intercept-only multilevel meta-regression: M3_preval)
pub.bias_preval_M3 <- rma.mv(yi = yi, V = V_preval,
  mods = ~  1 + sqrt_inv_n_tilda + year_c,
  random = list(~1 | study_id, ~1 | species_OTL, ~1 | obs), 
  R = list(species_OTL = phylo_cor_preval),
  data = preval_es,
  method = "REML", 
  test = "t") 

summary(pub.bias_preval_M3)

### Figure ###
## NOTE: Visual inspection is not recommended in datasets with multiple level of non-independence and high heterogeneity (Nakagawa et al. 2021)

# Raw residuals (observed - fitted)
res2 <- residuals(M3_preval, type = "response")

# Use original sampling SE for precision axis
sei2 <- sqrt(M3_preval$vi)

# Funnel of residuals vs precision (1/SE)
funnel(x = res2,
       sei = sei2,
       yaxis = "seinv",
       xlab  = "Residuals",
       ylab  = "Precision (1/SE)")
abline(v = 0, lty = 2)
```

### Prevalence ###
### Does baseline mortality rate (inherent mortality) affect effect size?
```{r}
## testing the moderating effect of baseline mortality rates on immunisation effectiveness
inherent_preval <- rma.mv(yi = yi, V = V_preval,
  mods = ~  1 + C_mortality,
  random = list(~1 | study_id, ~1 | species_OTL, ~1 | obs), 
  R = list(species_OTL = phylo_cor_preval),
  data = preval_es,
  method = "REML", 
  test = "t", control = list(optimizer = "bobyqa")) 

summary(inherent_preval) 
```

### Infection intensity ###
### Before meta-analysis => back transforming the log-scale value to the raw-scale value ###
```{r}
metadata <- read.csv("metadata.csv")

ln10 <- log(10)

metadata_new <- metadata %>%
  mutate(
    ## TREATMENT mean
    mean_T_trait = case_when(
      flag_unit_transformed == "FALSE" ~ mean_T_trait_initial,
      flag_unit_transformed == "ln(X)_NOSD" ~ exp(mean_T_trait_initial),
      flag_unit_transformed == "log10(X)" ~ {
        Z  <- mean_T_trait_initial * ln10
        SZ <- SD_T_initial * ln10
        exp(Z + (SZ^2)/2)
      },
      flag_unit_transformed == "log10(X+1)" ~ {
        Z  <- mean_T_trait_initial * ln10
        SZ <- SD_T_initial * ln10
        exp(Z + (SZ^2)/2) - 1
      },
      flag_unit_transformed == "log10(X+1)_NOSD" ~ {
        Z  <- mean_T_trait_initial * ln10
        exp(Z) - 1
      },
      TRUE ~ mean_T_trait_initial
    ),
    SD_T = case_when(
      flag_unit_transformed == "FALSE" ~ SD_T_initial,
      flag_unit_transformed == "log10(X)" ~ {
        Z  <- mean_T_trait_initial * ln10
        SZ <- SD_T_initial * ln10
        sqrt( (exp(SZ^2) - 1) * exp(2*Z + SZ^2) )
      },
      flag_unit_transformed == "log10(X+1)" ~ {
        ## subtracting 1 is a shift -> SD unchanged from the back-transformed X+1
        Z  <- mean_T_trait_initial * ln10
        SZ <- SD_T_initial * ln10
        sqrt( (exp(SZ^2) - 1) * exp(2*Z + SZ^2) )
      },
      TRUE ~ SD_T_initial
    ),
    mean_C_trait = case_when(
      flag_unit_transformed == "FALSE" ~ mean_C_trait_initial,
      flag_unit_transformed == "ln(X)_NOSD" ~ exp(mean_C_trait_initial),
      flag_unit_transformed == "log10(X)" ~ {
        Z  <- mean_C_trait_initial * ln10
        SZ <- SD_C_initial * ln10
        exp(Z + (SZ^2)/2)
      },
      flag_unit_transformed == "log10(X+1)" ~ {
        Z  <- mean_C_trait_initial * ln10
        SZ <- SD_C_initial * ln10
        exp(Z + (SZ^2)/2) - 1
      }, 
        flag_unit_transformed == "log10(X+1)_NOSD" ~ {
        Z  <- mean_C_trait_initial * ln10
        exp(Z) - 1
      },
      TRUE ~ mean_C_trait_initial
    ),
    SD_C = case_when(
      flag_unit_transformed == "FALSE" ~ SD_C_initial,
      flag_unit_transformed == "log10(X)" ~ {
        Z  <- mean_C_trait_initial * ln10
        SZ <- SD_C_initial * ln10
        sqrt( (exp(SZ^2) - 1) * exp(2*Z + SZ^2) )
      },
      flag_unit_transformed == "log10(X+1)" ~ {
        Z  <- mean_C_trait_initial * ln10
        SZ <- SD_C_initial * ln10
        sqrt( (exp(SZ^2) - 1) * exp(2*Z + SZ^2) )
      },
      TRUE ~ SD_C_initial
    )
  )
```

## infection intensity ##
## main analysis ##
## Removing dataset containing zero value (= uninfected individuals) ## 
```{R}
## Infection intensity should include only infected survivors.
## However, some contrasts included uninfected animals (Bd load = 0) 
## In these cases, the effect on infection intensity is conflated with that on prevalence 
## Here, we excluded those data sets to re-assess the effects.

## Remove contrasts that used a different definition to calculate pathogen load (k = 15)
intensity_INC <- metadata_new %>% 
  filter(Data_inclusiveness_flag %in% c("FALSE","MAYBE")) %>% 
  filter(trait == "intensity", 
         mean_T_trait > 0, mean_C_trait > 0, # ensure all values are positive (lnRR cannot be calculated with zero value)
         n_T_animals > 1, n_C_animals > 1) ## removing effect size with n = 1 (where SD cannot be estimated)

######################################################
## Using the 'missing-cases' method to calculate vi ##
######################################################

# Compute observed CV^2 (where available)
intensity_INC <- intensity_INC %>%
  mutate(CV2_T = (SD_T / mean_T_trait)^2,    # treatment CV^2
         CV2_C = (SD_C / mean_C_trait)^2)    # control CV^2

# Calculate the average between study CV, which will replace missing values
intensity_INC <- cv_avg(x = mean_T_trait, sd = SD_T, n = n_T_animals,
              group = study_id, label = "T", data = intensity_INC) #b_CV2_T (between-study treatment CV*2)
intensity_INC <- cv_avg(x = mean_C_trait, sd = SD_C, n = n_C_animals,
              group = study_id, label = "C", data = intensity_INC) #b_CV2_C (between-study control CV*2)

# Use weighted mean CV in replacement for where CV's are missing. Otherwise, calculate CV^2 of data that is known.
intensity_INC <- intensity_INC %>%
              mutate(cv2_T_new = if_else(is.na(CV2_T), b_CV2_T, CV2_T),
                     cv2_C_new = if_else(is.na(CV2_C), b_CV2_C, CV2_C))
    
# Calculate effect size (yi) and sampling variances (vi) using the 'missing-cases' method
intensity_data_INC <- intensity_INC %>%
  mutate(yi = lnrr_laj(m1 = mean_T_trait, m2 = mean_C_trait,
                    cv1_2 = cv2_T_new, cv2_2 = cv2_C_new,
                       n1 = n_T_animals, n2 = n_C_animals),
       vi = v_lnrr_laj(cv1_2 = cv2_T_new, n1 = n_T_animals,
                       cv2_2 = cv2_C_new, n2 = n_C_animals))

###########################################
## Non-independence: Shared animal group ##
###########################################
V <- vcalc(vi      = vi,
           cluster = study_exp_id,
           grp1    = T_ID, 
           grp2    = C_ID, 
           w1      = n_T_animals,
           w2      = n_C_animals,
           data = intensity_data_INC) 

######################################
## Random factor: observation-level ##
######################################

intensity_data_INC$obs <- 1:nrow(intensity_data_INC)

#################################
## Non-independence: phylogeny ##
#################################

# Species set from this filtered metadata
sp_data <- unique(trimws(gsub("\\s+", " ", intensity_data_INC$host_species)))
sp_data_treefmt <- gsub(" ", "_", sp_data)  # match typical tree tip style

# Prune the tree: only keep species present in BOTH tree and the filtered metadata
tips_to_keep <- intersect(tree$tip.label, sp_data_treefmt)
tree_pruned0 <- keep.tip(tree, tips_to_keep) 

# Force the tree to be binary
tree_pruned_INC <- multi2di(tree_pruned0)
is.binary (tree_pruned_INC) # yes, the tree is binary

# Create correlation phylogeny matrix for analysis
phylo_cor_INC <- vcv(tree_pruned_INC, cor = T)

# Create a column in metadata that matches the tree’s tip-label format
intensity_data_INC$species_OTL <- gsub(" ", "_", trimws(intensity_data_INC$host_species))

####################################
###  Overall immunisation effect ###
####################################

M3_INC <- rma.mv(yi = yi, V = V,
  mods = ~  1,
  random = list(~1 | study_id, ~1 | species_OTL, ~1 | obs),
  R = list(species_OTL = phylo_cor_INC),
  data = intensity_data_INC,
  method = "REML", test = "t")

summary (M3_INC)

## Estimate absolute heterogenesity
sum(M3_INC$sigma2)
## Estimate relative heterogeneity
orchaRd::i2_ml(M3_INC)
orchaRd::cvh2_ml(M3_INC)
orchaRd::m2_ml(M3_INC)

##########################
## Procedural variation ##
##########################

#### ** (1) Measurement effort throughout the experiment ##

# standardize (mean-centered) n_infection_measure
intensity_data_INC$n_infection_measure_c <- scale(intensity_data_INC$n_infection_measure, center = TRUE, scale = FALSE)

#### ** (2) Time elapsed between infection challenge and trait measurement ##

# standardize (mean-centered) average time at measure 
intensity_data_INC$time_measure   <- (intensity_data_INC$T_time_post + intensity_data_INC$C_time_post)/2
intensity_data_INC$time_measure_c <- scale(intensity_data_INC$time_measure, center = TRUE, scale = FALSE)

#### ** (3) Time interval between immunisation and infection challenge ##

# standardize (mean-centered) time intervals
intensity_data_INC$time_persist_c <- scale(intensity_data_INC$time_interval, center = TRUE, scale = FALSE)

#### MODEL accounting for (1) measurement effort (2) time interval between challenge and measurement and (3) time interval between immunisation and challenge
M3_covariate <- rma.mv(yi = yi, V = V,
  mods = ~ time_persist_c + time_measure_c + n_infection_measure_c,
  random = list(~1 | study_id, ~1 | species_OTL, ~1 | obs),
  R = list(species_OTL = phylo_cor),
  data = intensity_data_INC,
  method = "REML", test = "t")

summary (M3_covariate)

############################################
### Moderating effect: immunisation type ###
############################################

intensity_data_INC$treatment <- factor(
  intensity_data_INC$treatment,
  levels = c("other","synthetic_antiparastics", "dead_pathogen", "probiotic","live_pathogen","natural_chem"))


## Analysis: test for individual effect of each category
M_T_INC <- rma.mv(yi = yi, V = V,
  mods = ~  -1 + treatment,
  random = list(~1 | study_id, ~1 | species_OTL, ~1 | obs),
  R = list(species_OTL = phylo_cor_INC),
  data = intensity_data_INC,
  method = "REML", test = "t")

summary (M_T_INC) 

## Prediction intervals
intensity_data_INC$treatment <- droplevels(factor(intensity_data_INC$treatment))
levs <- levels(intensity_data_INC$treatment)

Xnew <- model.matrix(~ -1 + treatment,
                     data = data.frame(treatment = factor(levs, levels = levs)))

pred <- predict(M_T_INC, newmods = Xnew)

out_M_T_INC <- data.frame(treatment = levs, as.data.frame(pred))

out_M_T_INC[-1] <- round(out_M_T_INC[-1], 4)
out_M_T_INC

# Check k for each treatment type 
count <- intensity_data_INC %>% group_by(treatment) %>% summarise(total_k = n(),.groups = "drop") 
count

#####################################
### Moderating effect: life stage ###
#####################################

# Set as a factor
intensity_data_INC$life_stage_tested <- factor (intensity_data_INC$life_stage_tested, levels = c ("adult","juvenile", "larva"))

## Analysis: test for individual effect of each category
M_LS_INC <- rma.mv(yi = yi, V = V,
  mods = ~  -1 + life_stage_tested,
  random = list(~1 | study_id, ~1 | species_OTL, ~1 | obs),
  R = list(species_OTL = phylo_cor_INC),
  data = intensity_data_INC,
  method = "REML", test = "t")

summary (M_LS_INC) 

## Prediction intervals
levs <- levels(intensity_data_INC$life_stage_tested)
out_stage_int <- cbind(life_stage_tested = levs,
                   as.data.frame(predict(M_LS_INC,
                     newmods = model.matrix(~ -1 + life_stage_tested,
                       data = data.frame(life_stage_tested = levs)))))
options(digits = 4)
out_stage_int


# Check k per life stage
count <- intensity_data_INC %>%
  group_by(life_stage_tested) %>%
  summarise(total_k = n(),
            .groups = "drop")  
count

###################################
### Moderating effect: Taxonomy ###
###################################

intensity_data_INC$family <- factor(
  intensity_data_INC$family,
  levels = c("Bufonidae", "Ranidae", "Hylidae", "Alytidae", "Cryptobranchidae", "Sirenidae", "Plethodontidae", "Pipidae", "Ambystomatidae"))

## Analysis: test for individual effect of each category (taxonomaic families)
M_taxa_INC <- rma.mv(yi = yi, V = V,
  mods = ~  -1 + family,
  random = list(~1 | study_id, ~1 | species_OTL, ~1 | obs),
  R = list(species_OTL = phylo_cor_INC),
  data = intensity_data_INC,
  method = "REML", test = "t")

summary (M_taxa_INC) 

## Prediction intervals
intensity_data_INC$family <- droplevels(factor(intensity_data_INC$family))
levs <- levels(intensity_data_INC$family)
newdat <- data.frame(family = factor(levs, levels = levs))
Xnew <- model.matrix(~ -1 + family, data = newdat)
pred <- predict(M_taxa_INC, newmods = Xnew)
out_M_taxa_INC <- data.frame(family = levs, as.data.frame(pred))
out_M_taxa_INC[-1] <- round(out_M_taxa_INC[-1],4)
out_M_taxa_INC

# Check k for each family type
count <- intensity_data_INC %>% group_by(family) %>% summarise(total_k = n(),.groups = "drop") 
count

####################################
## Moderating effect: host origin ##
####################################

# Here I only consider "juvenile" while examining the effect of host origin (wild vs captive)

# Keep only juvenile rows
keepINC <- (intensity_data_INC$life_stage_tested %in% "juvenile") &
  (intensity_data_INC$host_origin %in% c("captive", "wild"))

# Subset the data
dat_INC <- intensity_data_INC[keepINC, ]

# Subset the already-built variance-covariance matrix
V_INC <- V [keepINC, keepINC]

# Re-generate the phylogeny matrix ###
tree <- read.tree("tree_for_imputation.tre")

# Species set from the new metadata
sp_data <- unique(trimws(gsub("\\s+", " ", dat_INC$host_species)))
sp_data_treefmt <- gsub(" ", "_", sp_data)  # match typical tree tip style

# Check the mismatch between species in my metadata and species in existing tree
sp_tree_spaces <- gsub("_", " ", tree$tip.label)
setdiff(sp_data, sp_tree_spaces) # In data but not in tree

# Prune the tree: only keep species present in BOTH tree and my metadata
tips_to_keep <- intersect(tree$tip.label, sp_data_treefmt)
tree_juv_INC <- keep.tip(tree, tips_to_keep) 

# Force the tree to be binary
tree_juv_INC <- multi2di(tree_juv_INC)
is.binary (tree_juv_INC) # yes, the tree is binary

# Create correlation phylogeny matrix for analysis
phylo_cor_juv_INC <- vcv(tree_juv_INC, cor = T)

# Create a column in metadata that matches the tree’s tip-label format
dat_INC $species_OTL <- gsub(" ", "_", trimws(dat_INC$host_species))

## Analysis: test for individual effect of each category

M_origin_juv_INC <- rma.mv(yi = yi, V = V_INC,
                       mods = ~ -1 + host_origin,
                       random = list(~1 | study_id, ~1 | species_OTL, ~1 | obs), 
                       R = list(species_OTL = phylo_cor_juv_INC), data = dat_INC,
                       method = "REML", test = "t")

summary(M_origin_juv_INC)  

## Prediction intervals
dat_INC$host_origin <- droplevels(factor(dat_INC$host_origin))
levs <- levels(dat_INC$host_origin)
newdat <- data.frame(host_origin = factor(levs, levels = levs))
Xnew <- model.matrix(~ -1 + host_origin, data = newdat)
pred <- predict(M_origin_juv_INC, newmods = Xnew)
out_host_origin_INC <- data.frame(host_origin = levs, as.data.frame(pred))
out_host_origin_INC[-1] <- round(out_host_origin_INC[-1], 4)
out_host_origin_INC

# Check K per juvenile origin
count <- dat_INC %>%
  group_by(host_origin) %>%
  summarise(total_k = n(),
            .groups = "drop")  
count
```
## infection intensity ##
## checking for publication bias ##
```{r}
# Standardize (mean-centered) publication year
intensity_data_INC$year_c <- scale(intensity_data_INC$publication_year, center = TRUE, scale = FALSE)

# Calculate the inverse of the "effective sample size" to account for unbalanced sampling lnRR 
intensity_data_INC$inv_n_tilda <-  with(intensity_data_INC, (n_T_animals + n_C_animals)/(n_T_animals*n_C_animals))
intensity_data_INC$sqrt_inv_n_tilda <-  with(intensity_data_INC, sqrt(inv_n_tilda))

pub_bias_M3_load <- rma.mv(yi = yi, V = V,
  mods = ~  1 + sqrt_inv_n_tilda + year_c,
  random = list(~1 | study_id, ~1 | species_OTL, ~1 | obs),
  R = list(species_OTL = phylo_cor_INC),
  data = intensity_data_INC,
  method = "REML",  test = "t") 

summary (pub_bias_M3_load)

### Funnel plots ###
## NOTE: Visual inspection is not recommended in dataset with multiple level of non-independence and high heterogeneity (Nakagawa et al. 2021)

# Raw residuals (observed - fitted)
res <- residuals(M3_INC, type = "response")

# Use original sampling SE for precision axis
sei <- sqrt(M3_INC$vi)

# Funnel of residuals vs precision (1/SE)
funnel(x = res,
       sei = sei,
       yaxis = "seinv",
       xlab  = "Residuals",
       ylab  = "Precision (1/SE)",
       xlim = c(-8.5,8.5))
abline(v = 0, lty = 2)

## output = 4 * 7 (portrait)
```

## infection intensity ##
## Do baseline mortality rate (= inherent susceptability) affect the effect size of infection loads?
```{r}
inherent_load <- rma.mv(yi = yi, V = V,
  mods = ~  1 + C_mortality,
  random = list(~1 | study_id, ~1 | species_OTL, ~1 | obs),
  R = list(species_OTL = phylo_cor_INC),
  data = intensity_data_INC,
  method = "REML",  test = "t") 

summary (inherent_load)
```

## infection intensity (log response ratio) ##
## sensitivity analysis - including zero value (uninfected individuals) dataset
```{r}
# Select intensity data
intensity <- metadata_new %>% 
  filter(trait == "intensity", 
         mean_T_trait > 0, mean_C_trait > 0, # ensure all values are positive (lnRR cannot be calculated with zero value)
         n_T_animals > 1, n_C_animals > 1) ## removing effect size with n = 1 (where SD cannot be estimated)

#################################
## Implementation of missing SD #
#################################

## Deal with observations with missing sampling variance (Cashin et al. 2013; Waddle PhD thesis, Becker et al. 2021; Barnett et al. 2023)
## code from Nakagawa et al 2022 (Ecology Letters) 
## missing cases method: https://alistairmcnairsenior.github.io/Miss_SD_Sim/

source("func.R") ## https://osf.io/sqr4w

# Compute observed CV^2 (where available)
intensity <- intensity %>%
  mutate(
    CV2_T = (SD_T / mean_T_trait)^2,   # treatment CV^2
    CV2_C = (SD_C / mean_C_trait)^2)    # control CV^2

# Calculate the average between study CV, which will replace missing values
intensity <- cv_avg(x = mean_T_trait, sd = SD_T, n = n_T_animals,
              group = study_id, label = "T", data = intensity) #b_CV2_T (between-study treatment CV*2)
intensity <- cv_avg(x = mean_C_trait, sd = SD_C, n = n_C_animals,
              group = study_id, label = "C", data = intensity) #b_CV2_C (between-study control CV*2)

# Use weighted mean CV in replacement for where CV's are missing. Otherwise, calculate CV^2 of data that is known.
intensity <- intensity %>%
              mutate(cv2_T_new = if_else(is.na(CV2_T), b_CV2_T, CV2_T),
                     cv2_C_new = if_else(is.na(CV2_C), b_CV2_C, CV2_C))
    
# Calculate effect size (yi) and sampling variances (vi) using missing-cases method
intensity_data <- intensity %>%
  mutate(
    yi = lnrr_laj(m1 = mean_T_trait, m2 = mean_C_trait,
                 cv1_2 = cv2_T_new, cv2_2 = cv2_C_new,
                 n1= n_T_animals, n2 = n_C_animals),
    vi = v_lnrr_laj(cv1_2 = cv2_T_new, n1 = n_T_animals,
                       cv2_2 = cv2_C_new, n2 = n_C_animals))

###########################################
## Non-independence: Shared animal group ##
###########################################

# Adjust non-independence of sampling variances/co-variances
V <- vcalc(vi      = vi,
           cluster = study_exp_id, # experiment level where shared value occurred
           grp1    = T_ID, # cohort ID of treatment
           grp2    = C_ID, # cohort ID of control
           w1      = n_T_animals,
           w2      = n_C_animals,
           data = intensity_data)

# Visualization 
corrplot::corrplot(cov2cor(V)) # I will replace "vi" using "V" in the following model

######################################
## Random factor: observation-level ##
######################################

intensity_data$obs <- 1:nrow(intensity_data)

##############################
## Random factor: phylogeny ##
##############################
tree <- read.tree("tree_for_imputation.tre")

# Species set from my metadata
sp_data <- unique(trimws(gsub("\\s+", " ", intensity_data$host_species)))
sp_data_treefmt <- gsub(" ", "_", sp_data)  # match typical tree tip style

# Check the mismatch between species in my metadata and species in existing tree
sp_tree_spaces <- gsub("_", " ", tree$tip.label)
setdiff(sp_data, sp_tree_spaces) # In data but not in tree

# Prune the tree: only keep species present in BOTH tree and my metadata
tips_to_keep <- intersect(tree$tip.label, sp_data_treefmt)
tree_pruned0 <- keep.tip(tree, tips_to_keep) 

# Force the tree to be binary
tree_pruned <- multi2di(tree_pruned0)
is.binary (tree_pruned) # yes, the tree is binary

# Create correlation phylogeny matrix for analysis
phylo_cor <- vcv(tree_pruned, cor = T)

# Create a column in metadata that matches the tree’s tip-label format
intensity_data$species_OTL <- gsub(" ", "_", trimws(intensity_data$host_species))

####################
## Overall effect ##
####################

M3_load <- rma.mv(yi = yi, V = V,
  mods = ~  1,
  random = list(~1 | study_id, ~1 | species_OTL, ~1 | obs),
  R = list(species_OTL = phylo_cor),
  data = intensity_data,
  method = "REML",  test = "t") 
  
summary (M3_load) # final model for overall treatment effect on pathogen load
predict(M3_load)
## Estimate absolute heterogenesity
sum(M3_load$sigma2)
## Estimate relative heterogeneity
orchaRd::i2_ml(M3_load)
orchaRd::cvh2_ml(M3_load)
orchaRd::m2_ml(M3_load)

#############################################
## categorical variable: immunisation type ##
#############################################

# Set as a factor
intensity_data$treatment <- factor(
  intensity_data$treatment,
  levels = c("other", "probiotic", "synthetic_antiparastics","natural_chem", "dead_pathogen", "live_pathogen"))

## Analysis
M_treatment <- rma.mv(yi = yi, V = V,
  mods = ~  -1 + treatment,
  random = list(~1 | study_id, ~1 | species_OTL, ~1 | obs),
  R = list(species_OTL = phylo_cor),
  data = intensity_data,
  method = "REML", test = "t")

summary (M_treatment) 

## Prediction intervals
levs <- levels(intensity_data$treatment)
out_treat_int <- cbind(treatment = levs,
                   as.data.frame(predict(M_treatment,
                     newmods = model.matrix(~ -1 + treatment,
                       data = data.frame(treatment = levs)))))
options(digits = 4)
out_treat_int

# Count k per treatment type
count <- intensity_data %>%
  group_by(treatment) %>%
  summarise(total_k = n(),
            .groups = "drop")  
count

######################################
## categorical variable: life stage ##
######################################

# Set as factor
intensity_data$life_stage_tested <- factor(
  intensity_data$life_stage_tested,
  levels = c("larva","juvenile","adult"))

## ANALYSIS: test for the effect for individual life stage
M_stage <- rma.mv(yi = yi, V = V,
  mods = ~  -1 + life_stage_tested,
  random = list(~1 | study_id, ~1 | species_OTL, ~1 | obs),
  R = list(species_OTL = phylo_cor),
  data = intensity_data,
  method = "REML", test = "t")

summary (M_stage) 

## Prediction intervals
levs <- levels(intensity_data$life_stage_tested)
out_stage_int <- cbind(life_stage_tested = levs,
                   as.data.frame(predict(M_stage,
                     newmods = model.matrix(~ -1 + life_stage_tested,
                       data = data.frame(life_stage_tested = levs)))))
options(digits = 4)
out_stage_int


# Count k per life stage
intensity_data %>%
  group_by(life_stage_tested) %>%
  summarise(total_n = n(),
            .groups = "drop") 

####################################
## categorical variable: taxonomy ##
####################################

# Set as a factor
intensity_data$family <- factor(intensity_data$family)

intensity_data$family <- factor(
  intensity_data$family,
  levels = c("Bufonidae", "Ranidae", "Hylidae", "Alytidae", "Cryptobranchidae", "Sirenidae", "Plethodontidae", "Pipidae", "Ambystomatidae"))

## ANALYSIS: test for the effect for individual family
M_taxa <- rma.mv(yi = yi, V = V,
  mods = ~  -1 + family,
  random = list(~1 | study_id, ~1 | species_OTL, ~1 | obs),
  R = list(species_OTL = phylo_cor),
  data = intensity_data,
  method = "REML", test = "t")

summary (M_taxa) 

## Prediction intervals
levs <- gsub("^family", "", names(coef(M_taxa)))

Xnew <- diag(length(coef(M_taxa)))
colnames(Xnew) <- names(coef(M_taxa))

pred_taxa <- predict(M_taxa, newmods = Xnew)
options(digits = 4)
out_taxa <- data.frame(family = levs, as.data.frame(pred_taxa))
out_taxa

# Count k per family type
count <- intensity_data %>%
  group_by(family) %>%
  summarise(total_k = n(),
            .groups = "drop")  
count

#######################################
## categorical variable: host origin ##
#######################################

# Count origin per life stage
origin_count <- intensity_data %>%
  group_by(life_stage_tested, host_origin) %>%
  summarise(total_n = n(),
            .groups = "drop") 
origin_count

# most adults (15/16 = 94%) and all larvae were from the wild, whereas juveniles originated from either captive or wild populations. 
# I therefore subdivided the dataset to test the effect of origin in juveniles only. 

# Keep only juvenile rows
keep2 <- (intensity_data$life_stage_tested %in% "juvenile") &
  (intensity_data$host_origin %in% c("wild", "captive"))

# Subset the data
dat1in <- intensity_data[keep2, ]

# Subset the already-built variance-covariance matrix
V_keep2 <- V [keep2, keep2]

# Re-generate the phylogeny matrix ###
tree <- read.tree("tree_for_imputation.tre")

# Species set from the new metadata
sp_data <- unique(trimws(gsub("\\s+", " ", dat1in$host_species)))
sp_data_treefmt <- gsub(" ", "_", sp_data)  # match typical tree tip style

# Check the mismatch between species in my metadata and species in existing tree
sp_tree_spaces <- gsub("_", " ", tree$tip.label)
setdiff(sp_data, sp_tree_spaces) # In data but not in tree

# Prune the tree: only keep species present in BOTH tree and my metadata
tips_to_keep <- intersect(tree$tip.label, sp_data_treefmt)
tree_pruned_juv <- keep.tip(tree, tips_to_keep) 

# Force the tree to be binary
tree_pruned_juv <- multi2di(tree_pruned_juv)
is.binary (tree_pruned_juv) # yes, the tree is binary

# Create correlation phylogeny matrix for analysis
phylo_cor_juv <- vcv(tree_pruned_juv, cor = T)

# Create a column in metadata that matches the tree’s tip-label format
dat1in$species_OTL <- gsub(" ", "_", trimws(dat1in$host_species))

# Set as a factor
dat1in$host_origin <- factor(
  dat1in$host_origin,
  levels = c("captive", "wild"))

## ANALYSIS: test for individal effect of origin
M_juv_origin <- rma.mv(yi = yi, V = V_keep2,
                       mods = ~ -1 + host_origin,
                       random = list(~1 | study_id, ~1 | species_OTL, ~1 | obs), 
                       R = list(species_OTL = phylo_cor_juv), data = dat1in,
                       method = "REML", test = "t")

summary(M_juv_origin) 

## Prediction intervals
levs <- levels(dat1in$host_origin)
out_origin_int <- cbind(host_origin = levs,
                   as.data.frame(predict(M_juv_origin,
                     newmods = model.matrix(~ -1 + host_origin,
                       data = data.frame(host_origin = levs)))))
options(digits = 4)
out_origin_int

# Count k per origin
dat1in %>%
  group_by(host_origin) %>%
  summarise(total_n = n(),
            .groups = "drop") 

```

## infection intensity  ##
## sensitivity analysis (II) ##
## Removing contrasts with log-transformed values ## 
```{r}
## Here, I exclude records with back-transformed data to check whether the effects change.

## Remove contrasts (k = 18) with transformed values
intensity_NTV <- metadata_new %>% 
  filter(flag_unit_transformed == "FALSE") %>% 
  filter(Data_inclusiveness_flag %in% c("FALSE","MAYBE")) %>% 
  filter(trait == "intensity", 
         mean_T_trait > 0, mean_C_trait > 0, # ensure all values are positive (lnRR cannot be calculated with zero value)
         n_T_animals > 1, n_C_animals > 1) ## removing effect size with n = 1 (where SD cannot be estimated)

##################################################################
## using the 'missing-cases' method to calculate unavailable vi ##
##################################################################

# Compute observed CV^2 (where available)
intensity_NTV <- intensity_NTV %>%
  mutate(CV2_T = (SD_T / mean_T_trait)^2,    # treatment CV^2
         CV2_C = (SD_C / mean_C_trait)^2)    # control CV^2

# Calculate the average between-study CV, which will replace missing values
intensity_NTV <- cv_avg(x = mean_T_trait, sd = SD_T, n = n_T_animals,
              group = study_id, label = "T", data = intensity_NTV) #b_CV2_T (between-study treatment CV*2)
intensity_NTV <- cv_avg(x = mean_C_trait, sd = SD_C, n = n_C_animals,
              group = study_id, label = "C", data = intensity_NTV) #b_CV2_C (between-study control CV*2)

# Use weighted mean CV in replacement for where CV's are missing. Otherwise, calculate CV^2 of data that is known.
intensity_NTV <- intensity_NTV %>%
              mutate(cv2_T_new = if_else(is.na(CV2_T), b_CV2_T, CV2_T),
                     cv2_C_new = if_else(is.na(CV2_C), b_CV2_C, CV2_C))
    
# Calculate effect size (yi) and sampling variances (vi) using the 'missing-cases' method
intensity_data_NTV <- intensity_NTV %>%
  mutate(
    yi = lnrr_laj(m1 = mean_T_trait, m2 = mean_C_trait,
                 cv1_2 = cv2_T_new, cv2_2 = cv2_C_new,
                 n1= n_T_animals, n2 = n_C_animals),
    vi = v_lnrr_laj (cv1_2 = cv2_T_new, n1 = n_T_animals,
                       cv2_2 = cv2_C_new, n2 = n_C_animals))

###########################################
## Non-independence: Shared animal group ##
###########################################
V <- vcalc(vi      = vi,
           cluster = study_exp_id, 
           grp1    = T_ID, 
           grp2    = C_ID, 
           w1      = n_T_animals,
           w2      = n_C_animals,
           data = intensity_data_NTV) ## matrix

######################################
## Random factor: observation-level ##
######################################

intensity_data_NTV$obs <- 1:nrow(intensity_data_NTV)

#################################
## Non-independence: phylogeny ##
#################################

# Species set from this filtered metadata
sp_data <- unique(trimws(gsub("\\s+", " ", intensity_data_NTV$host_species)))
sp_data_treefmt <- gsub(" ", "_", sp_data)  # match typical tree tip style

# Prune the tree: only keep species present in BOTH tree and the filtered metadata
tips_to_keep <- intersect(tree$tip.label, sp_data_treefmt)
tree_pruned0 <- keep.tip(tree, tips_to_keep) 

# Force the tree to be binary
tree_pruned_NTV <- multi2di(tree_pruned0)
is.binary (tree_pruned_NTV) # yes, the tree is binary

# Create correlation phylogeny matrix for analysis
phylo_cor_NTV <- vcv(tree_pruned_NTV, cor = T)

# Create a column in metadata that matches the tree’s tip-label format
intensity_data_NTV$species_OTL <- gsub(" ", "_", trimws(intensity_data_NTV$host_species))

#########################################
### Moderating effect: Overall effect ###
#########################################

M3_NTV <- rma.mv(yi = yi, V = V,
  mods = ~  1,
  random = list(~1 | study_id, ~1 | species_OTL, ~1 | obs),
  R = list(species_OTL = phylo_cor_NTV),
  data = intensity_data_NTV,
  method = "REML", test = "t")

summary (M3_NTV) 
predict (M3_NTV)

###########################################
### Moderating effect: immunisation type ##
###########################################

# Set as a factor
intensity_data_NTV$treatment <- factor(
  intensity_data_NTV$treatment,
  levels = c("other", "probiotic", "synthetic_antiparastics","natural_chem", "dead_pathogen", "live_pathogen"))

## Analysis: test for individual effect of treatment type
M_T_NTV <- rma.mv(yi = yi, V = V,
  mods = ~  -1 + treatment,
  random = list(~1 | study_id, ~1 | species_OTL, ~1 | obs),
  R = list(species_OTL = phylo_cor_NTV),
  data = intensity_data_NTV,
  method = "REML", test = "t")

summary (M_T_NTV) 

## Prediction intervals
intensity_data_NTV$treatment <- droplevels(factor(intensity_data_NTV$treatment))
levs <- levels(intensity_data_NTV$treatment)
newdat <- data.frame(treatment = factor(levs, levels = levs))
Xnew <- model.matrix(~ -1 + treatment, data = newdat)
pred <- predict(M_T_NTV, newmods = Xnew)
out_M_T_NTV <- data.frame(treatment = levs, as.data.frame(pred))
out_M_T_NTV[-1] <- round(out_M_T_NTV[-1], 4)
out_M_T_NTV

# Check k per treatment type after excluding the transformed value
count <- intensity_data_NTV %>% group_by(treatment) %>% summarise(total_k = n(),.groups = "drop") 
count

#####################################
### Moderating effect: life stage ###
#####################################

# Set as a factor
intensity_data_NTV$life_stage_tested <- as.factor (intensity_data_NTV$life_stage_tested)

# ANALYSIS: test for individual effect on life stage
M_L_NTV <- rma.mv(yi = yi, V = V,
                  mods = ~  -1 + life_stage_tested,
                  random = list(~1 | study_id, ~ 1 | species_OTL, ~1 | obs),
                  R = list(species_OTL = phylo_cor_NTV),
                  data = intensity_data_NTV,
                  method = "REML", test = "t")

summary (M_L_NTV) 

## Prediction intervals
intensity_data_NTV$life_stage_tested <- droplevels(factor(intensity_data_NTV$life_stage_tested))
levs <- levels(intensity_data_NTV$life_stage_tested)
newdat <- data.frame(life_stage_tested = factor(levs, levels = levs))
Xnew <- model.matrix(~ -1 + life_stage_tested, data = newdat)
pred <- predict(M_L_NTV, newmods = Xnew)
out_M_L_NTV <- data.frame(life_stage_tested = levs, as.data.frame(pred))
out_M_L_NTV[-1] <- round(out_M_L_NTV[-1], 4)
out_M_L_NTV

# Check k per life stage after excluding the transformed value
count <- intensity_data_NTV %>%
  group_by(life_stage_tested) %>%
  summarise(total_k = n(),
            .groups = "drop")  
count

###################################
### Moderating effect: taxonomy ###
###################################

intensity_data_NTV$family <- as.factor (intensity_data_NTV$family)

## Analysis: test for individual effect of treatment type
M_Taxa_NTV <- rma.mv(yi = yi, V = V,
  mods = ~  -1 + family,
  random = list(~1 | study_id, ~1 | species_OTL, ~1 | obs),
  R = list(species_OTL = phylo_cor_NTV),
  data = intensity_data_NTV,
  method = "REML", test = "t")

summary (M_Taxa_NTV) 

## Prediction intervals
intensity_data_NTV$family <- droplevels(factor(intensity_data_NTV$family))
levs <- levels(intensity_data_NTV$family)
newdat <- data.frame(family = factor(levs, levels = levs))
Xnew <- model.matrix(~ -1 + family, data = newdat)
pred <- predict(M_Taxa_NTV, newmods = Xnew)
out_M_Taxa_NTV <- data.frame(family = levs, as.data.frame(pred))
out_M_Taxa_NTV[-1] <- round(out_M_Taxa_NTV[-1], 4)
out_M_Taxa_NTV

# Check k per treatment type after excluding the transformed value
count <- intensity_data_NTV %>% group_by(family) %>% summarise(total_k = n(),.groups = "drop") 
count

####################################
## Moderating effect: host origin ##
####################################

# Here I only consider "juvenile" while examining the effect of host origin (wild vs captive)

# Keep only juvenile rows
keepNTV <- (intensity_data_NTV$life_stage_tested %in% "juvenile") &
  (intensity_data_NTV$host_origin %in% c("captive", "wild"))

# Subset the data
dat_NTV <- intensity_data_NTV[keepNTV, ]

# Subset the already-built variance-covariance matrix
V_NTV <- V [keepNTV, keepNTV]

# Re-generate the phylogeny matrix ###
tree <- read.tree("tree_for_imputation.tre")

# Species set from the new metadata
sp_data <- unique(trimws(gsub("\\s+", " ", dat_NTV$host_species)))
sp_data_treefmt <- gsub(" ", "_", sp_data)  # match typical tree tip style

# Check the mismatch between species in my metadata and species in existing tree
sp_tree_spaces <- gsub("_", " ", tree$tip.label)
setdiff(sp_data, sp_tree_spaces) # In data but not in tree

# Prune the tree: only keep species present in BOTH tree and my metadata
tips_to_keep <- intersect(tree$tip.label, sp_data_treefmt)
tree_juv_NTV <- keep.tip(tree, tips_to_keep) 

# Force the tree to be binary
tree_juv_NTV <- multi2di(tree_juv_NTV)
is.binary (tree_juv_NTV) # yes, the tree is binary

# Create correlation phylogeny matrix for analysis
phylo_cor_juv_NTV <- vcv(tree_juv_NTV, cor = T)

# Create a column in metadata that matches the tree’s tip-label format
dat_NTV$species_OTL <- gsub(" ", "_", trimws(dat_NTV$host_species))

## Run model

## Analysis: test for inidivdual effect on juvenile origin
M_origin_juv_NTV <- rma.mv(yi = yi, V = V_NTV,
                       mods = ~ -1 + host_origin,
                       random = list(~1 | study_id, ~1 | species_OTL, ~1 | obs), 
                       R = list(species_OTL = phylo_cor_juv_NTV), data = dat_NTV,
                       method = "REML", test = "t")

summary(M_origin_juv_NTV) 

## Prediction intervals
dat_NTV$host_origin <- droplevels(factor(dat_NTV$host_origin))
levs <- levels(dat_NTV$host_origin)
newdat <- data.frame(host_origin= factor(levs, levels = levs))
Xnew <- model.matrix(~ -1 + host_origin, data = newdat)
pred <- predict(M_origin_juv_NTV, newmods = Xnew)
out_M_origin_juv_NTV <- data.frame(host_origin = levs, as.data.frame(pred))
out_M_origin_juv_NTV[-1] <- round(out_M_origin_juv_NTV[-1], 4)
out_M_origin_juv_NTV

# Count k per origin for juveniles only

dat_NTV %>%
  group_by(host_origin) %>%
  summarise(total_n = n(),
            .groups = "drop") 
            
```

## mortality (log risk ratio) ##
## main analysis
```{r}
## Read.table
metadata <- read.csv("metadata.csv")
mortal <- metadata %>% filter(trait == "mortality")

## Create new columns
mortal$T_deaths <- round(mortal$mean_T_trait/100 * mortal$n_T_animals)
mortal$T_survivors <- round(mortal$n_T_animals - mortal$T_deaths)

mortal$C_deaths <- round(mortal$mean_C_trait/100 * mortal$n_C_animals)
mortal$C_survivors <- round(mortal$n_C_animals - mortal$C_deaths)

mortal <- mortal %>% filter(!(T_deaths == 0 & C_deaths == 0)) ## remove column when no deaths at all        

#############################
## log risk ratio (log RR) ##
#############################

mort_es <- escalc(
  measure = "RR",
  ai = T_deaths,  bi = T_survivors, # ai = event, bi = non-event
  ci = C_deaths,  di = C_survivors, # ci = event, di = non-event
  data = mortal,
  add = 0.5, to = "only0") ## add 0.5 when a cell is zero 

###########################################
## Non-independence: Shared animal group ##
###########################################

# Adjust non-independence of sampling variances/co-variances
V_mor <- vcalc(vi      = vi,
           cluster = study_exp_id, # experiment level where shared value occurred
           grp1    = T_ID, # cohort ID of treatment
           grp2    = C_ID, # cohort ID of control
           w1      = n_T_animals,
           w2      = n_C_animals,
           data = mort_es)

# Visualize
corrplot::corrplot(cov2cor(V_mor)) 

######################################
## Random factor: observation-level ##
######################################

mort_es$obs <- 1:nrow(mort_es)

##############################
## Random factor: phylogeny ##
##############################

# Read the tree from Pottier et al. (2024) Nature:
# https://github.com/p-pottier/Vulnerability_amphibians_global_warming 

tree <- read.tree("tree_for_imputation.tre")

# Species set from the metadata
sp_data <- mort_es$host_species %>% as.character() %>% str_squish() %>% str_trim()

# Match typical tree tip style
sp_data_treefmt <- gsub(" ", "_", sp_data)  

# Check any mismatch between species in the metadata and species in existing tree
sp_tree_spaces <- gsub("_", " ", tree$tip.label)
setdiff(unique(sp_data), sp_tree_spaces) # In data but not in tree

# Prune the tree: only keep species present in BOTH tree and my metadata
tips_to_keep <- intersect(tree$tip.label, unique(sp_data_treefmt))
tree_pruned_mor <- keep.tip(tree, tips_to_keep) 

# Force the tree to be binary
tree_pruned_mor <- multi2di(tree_pruned_mor)
is.binary (tree_pruned_mor) # yes, the tree is binary

# Create correlation phylogeny matrix for analysis
phylo_cor_mor <- vcv(tree_pruned_mor, corr = TRUE)

# Create a column in metadata that matches the tree’s tip-label format
mort_es <- mort_es %>% mutate(species_OTL = gsub(" ", "_", str_squish(as.character(host_species))), species_OTL = factor(species_OTL))

####################
## Overall effect ##
####################

M3_mor <- rma.mv(yi = yi, V = V_mor,
  mods = ~  1,
  random = list(~1 | study_id, ~1 | species_OTL, ~1 | obs), 
  R = list(species_OTL = phylo_cor_mor),
  data = mort_es,
  method = "REML", 
  test = "t",
  control = list(optimizer = "optim", optmethod = "BFGS", maxit = 20000, reltol = 1e-10)) 

summary(M3_mor) # final model for overall treatment effect on pathogen load

## Heterogeneity
orchaRd::i2_ml(M3_mor)
orchaRd::cvh2_ml(M3_mor)
orchaRd::m2_ml(M3_mor)


##########################
## Procedural variation ##
##########################

#### (1) Time elapsed between infection challenge and trait measurement ###

# Assume higher mortality risk in controls than in treated groups when mortality is measured long after Bd exposure
# (i.e., more time for disease to develop, so more control individuals have died).
# A one-day increase in elapsed time → negative log RR (treatment vs control).

# Time elapsed has been extracted separately for the control and treatment groups. Two values are highly correlated (because both groups were often measured together) and cannot be treated as two separate moderators in a model (an issue of collinarity). I took the average time_elapsed.

# standardize (mean-centered) average time at measure 
mort_es$avg_time_measure   <- (mort_es$T_time_post + mort_es$C_time_post)/2
mort_es$time_measure_c <- scale(mort_es$avg_time_measure, center = TRUE, scale = FALSE)

#### (2) Time interval between immunisation and pathogen challenge ###

# Assuming increased mortality risk in treated groups when pathogen re-exposure occurs long after completion of treatment 
# (i.e., treatment-induced protection wanes over time)
# A one-day increase in the time interval → positive log RR (treatment vs control).

# standardize (mean-centered) time intervals
mort_es$time_persist_c <- scale(mort_es$time_interval, center = TRUE, scale = FALSE)

# MODEL accounting for (1) time interval between infection challenge and outcome measurement and (2) time interval between immunisation and infection challenge
M3_cov <- rma.mv(yi = yi, V = V_mor,
  mods = ~  1 + time_persist_c + time_measure_c,
  random = list(~1 | study_id, ~1 | species_OTL, ~1 | obs), 
  R = list(species_OTL = phylo_cor_mor),
  data = mort_es,
  method = "REML", 
  test = "t",
  control = list(optimizer = "optim", optmethod = "BFGS", maxit = 20000, reltol = 1e-10)) 

summary(M3_cov) ## both were non-significant

#############################################
## categorical variable: immunisation type ##
#############################################

mort_es$treatment <- factor(
  mort_es$treatment,
  levels = c("synthetic_antiparastics", "dead_pathogen", "probiotic","live_pathogen","natural_chem"))

## ANALYSIS: test for the effect for individual treatment

M_treat_dead1 <- rma.mv(yi = yi, V = V_mor,
  mods = ~  -1 + treatment,
  random = list(~1 | study_id, ~1 | species_OTL, ~1 | obs), 
  R = list(species_OTL = phylo_cor_mor), data = mort_es,
  method = "REML", test = "t",
  control = list(optimizer = "optim", optmethod = "BFGS", maxit = 20000, reltol = 1e-10)) 

summary(M_treat_dead1) 

## Prediction intervals
levs <- levels(mort_es$treatment)
out_mor <- cbind(treatment = levs,
                   as.data.frame(predict(M_treat_dead1,
                     newmods = model.matrix(~ -1 + treatment,
                       data = data.frame(treatment = levs)))))
options(digits = 3)
out_mor

# Check k per treatment type
count <- mort_es %>%
  group_by(treatment) %>%
  summarise(total_k = n(),
            .groups = "drop")  
count

######################################
## categorical variable: life stage ##
######################################

# Set as a factor
mort_es$life_stage_tested <- factor(
  mort_es$life_stage_tested,
  levels = c("adult","juvenile", "larva"))

## ANALYSIS: test for the effect for individual life stage
M_stage_dead <- rma.mv(yi = yi, V = V_mor,
  mods = ~  -1 + life_stage_tested,
  random = list(~1 | study_id, ~1 | species_OTL, ~1 | obs), 
  R = list(species_OTL = phylo_cor_mor), data = mort_es,
  method = "REML", test = "t",
  control = list(optimizer = "optim", optmethod = "BFGS", maxit = 20000, reltol = 1e-10)) 

summary(M_stage_dead) 

## Prediction intervals
levs <- levels(mort_es$life_stage_tested)
out_mor_stage <- cbind(life_stage_tested = levs,
                   as.data.frame(predict(M_stage_dead,
                     newmods = model.matrix(~ -1 + life_stage_tested,
                       data = data.frame(life_stage_tested = levs)))))
options(digits = 3)
out_mor_stage

# Count the k per life stage
count <- mort_es %>%
  group_by(life_stage_tested) %>%
  summarise(total_k = n(),
            .groups = "drop")  
count

#######################################
## categorical variable: host origin ##
#######################################

origin_count <- mort_es %>%
  group_by(host_origin, life_stage_tested) %>%
  summarise(total_n = n(),
            .groups = "drop") 

# 'Larvae' were only sourced from wild, so I removed "larva" group here and tested the effect of host origin (wild vs captive) for "adults" and "juveniles" only.

# Keep only adult + juvenile rows
keep1 <- mort_es$life_stage_tested %in% c("adult","juvenile")

# Subset the already-built variance-covariance matrix
V_keep1 <- V_mor[keep1, keep1]

# Subset the data (random effects & phylo R will align by factor levels automatically)
dat1 <- mort_es[keep1, ]

# Set as a factor
dat1$host_origin <- factor(
  dat1$host_origin,
  levels = c("captive", "wild"))

# Analysis: test for individual moderating effect of origin
M_host_dead1 <- rma.mv(yi = yi, V = V_keep1,
                         mods = ~ -1 + host_origin,
                         random = list(~1 | study_id, ~1 | species_OTL, ~1 | obs), 
                         R = list(species_OTL = phylo_cor_mor), data = dat1,
                         method = "REML", test = "t",
                         control = list(optimizer = "optim", optmethod = "BFGS", maxit = 20000, reltol = 1e-10)) 

summary(M_host_dead1) 

## Prediction intervals
levs <- levels(dat1$host_origin)
out_mor_origin <- cbind(host_origin = levs,
                   as.data.frame(predict(M_host_dead1,
                     newmods = model.matrix(~ -1 + host_origin,
                       data = data.frame(host_origin = levs)))))
options(digits = 3)
out_mor_origin

# Count the k per origin
count <- dat1 %>%
  group_by(host_origin) %>%
  summarise(total_k = n(),
            .groups = "drop")  
count

#########################################
## categorical variable: Host taxonomy ##
#########################################
mort_es$family <- factor(
  mort_es$family,
  levels = c("Bufonidae", "Ranidae", "Hylidae", "Alytidae", "Cryptobranchidae", "Sirenidae", "Plethodontidae", "Pipidae", "Ambystomatidae"))

## ANALYSIS: Individual effect on ech host taxonomy

M_taxa_dead1 <- rma.mv(yi = yi, V = V_mor,
  mods = ~  -1 + family,
  random = list(~1 | study_id, ~1 | species_OTL, ~1 | obs), 
  R = list(species_OTL = phylo_cor_mor), data = mort_es,
  method = "REML", test = "t",
  control = list(optimizer = "optim", optmethod = "BFGS", maxit = 20000, reltol = 1e-10)) 

summary(M_taxa_dead1) 

## Prediction intervals
levs <- gsub("^family", "", names(coef(M_taxa_dead1)))  # nice labels (optional)

Xnew <- diag(length(coef(M_taxa_dead1)))
colnames(Xnew) <- names(coef(M_taxa_dead1))

pred_taxa <- predict(M_taxa_dead1, newmods = Xnew)
options(digits = 4)
out_taxa <- data.frame(family = levs, as.data.frame(pred_taxa))
out_taxa

# Check k per family type
count <- mort_es %>%
  group_by(family) %>%
  summarise(total_k = n(),
            .groups = "drop")  
count
```

## mortality ##
## checking for publication bias ##
``` {r}
## Nakagawa et al (2021) Methods in Ecology and Evolution
## file:///Users/s449735/Documents/GitHub/amphib_chytrid/STEP%204%20-%20analysis/mee313724-sup-0001-Supinfo.html

# Standardize (mean-centered) publication year (to test for time-lag bias)
mort_es$year_c <- scale(mort_es$publication_year, center = TRUE, scale = FALSE)

# Calculate the inverse of the "effective sample size" (to test for small-study effects)
mort_es$inv_n_tilda <-  with(mort_es, (n_T_animals + n_C_animals)/(n_T_animals*n_C_animals))
mort_es$sqrt_inv_n_tilda <-  with(mort_es, sqrt(inv_n_tilda))

# Add sqrt__inverse_effective_sample_size to the meta-regression: M3_mor
pub_bias_M3_mor <- rma.mv(yi = yi, V = V_mor,
  mods = ~  1 + sqrt_inv_n_tilda + year_c,
  random = list(~1 | study_id, ~1 | species_OTL, ~1 | obs), 
  R = list(species_OTL = phylo_cor_mor),
  data = mort_es,
  method = "REML", 
  test = "t",
  control = list(optimizer = "optim", optmethod = "BFGS", maxit = 20000, reltol = 1e-10)) 

summary(pub_bias_M3_mor)

### Funnel plots ###

## NOTE: Visual inspection is not recommended in dataset with multiple level of non-independence and high heterogeneity (Nakagawa et al. 2021)

# Raw residuals (observed - fitted)
res <- residuals(M3_mor, type = "response")

# Use original sampling SE for precision axis
sei <- sqrt(M3_mor$vi)

# Funnel of residuals vs precision (1/SE)
funnel(x = res,
       sei = sei,
       yaxis = "seinv",
       xlab  = "Residuals",
       ylab  = "Precision (1/SE)")
abline(v = 0, lty = 2)

## output = 4 * 7 (portrait)
```

### mortality ###
## Do baseline mrotality rates affect effect size of mortality? 
```{R}
inherent_mor <- rma.mv(yi = yi, V = V_mor,
  mods = ~  1 + C_mortality,
  random = list(~1 | study_id, ~1 | species_OTL, ~1 | obs), 
  R = list(species_OTL = phylo_cor_mor),
  data = mort_es,
  method = "REML", 
  test = "t",
  control = list(optimizer = "optim", optmethod = "BFGS", maxit = 20000, reltol = 1e-10)) 

summary(inherent_mor)
```

### Infection progression ###
## Within-experiment correlation between effect sizes of three infection outcomes
```{r}
# https://www.metafor-project.org/doku.php/tips:i2_multilevel_multivariate

####################
## pool datasets ##
####################

# select columns from datasets of each infection outcome that are needed for multivariate models
mort_es1 <- mort_es %>% select(trait, yi, vi, study_id, study_exp_id, T_cohort_ID, host_species, family, treatment, life_stage_tested, host_origin, T_ID, C_ID, n_T_animals, n_C_animals)

preval_es1 <- preval_es %>% select(trait, yi, vi, study_id, study_exp_id, T_cohort_ID, host_species, family, treatment, life_stage_tested, host_origin, T_ID, C_ID, n_T_animals, n_C_animals)

int_es1 <- intensity_data_INC %>% select(trait, yi, vi, study_id, study_exp_id, T_cohort_ID, host_species, family, treatment, life_stage_tested, host_origin, T_ID, C_ID, n_T_animals, n_C_animals)

# pool three datasets
all_es <- bind_rows (mort_es1, preval_es1,int_es1)

# Create a new column called "cohort_id", which pools the three measurements (prevalence, infection load, mortality) from the same animal cohort.
# Note that this differs from T_ID and C_ID, which identify value of a single trait repeatedly used across different effect comparisons (e.g., shared controls).
all_es <- all_es %>%
  mutate(cohort_id = paste0(study_exp_id, sep = "_", T_cohort_ID))

########################
## Effect correlation ##
########################

# Model: within-EXPERIMENT correlation between effect sizes of three outcomes (cannot be estimated)
exp_cor <- rma.mv(yi = yi, V = vi,
                     mods = ~  -1 + trait,
                     random = list(~ trait -1 | study_exp_id),
                     struc = "UN",
                     data = all_es)


summary (exp_cor)
confint (exp_cor)

```
            
