---
title: "Meta-analysis_plot"
output: html_document
date: "2025-09-30"
---
## load package
```{r}
library(dplyr)
library(tidyr)   
library(ggplot2)
library(sf)

install.packages("remotes")
install.packages("rnaturalearth")

# Then install orchaRd from the developers
remotes::install_github("daniel1noble/orchaRd")
library(orchaRd)

# --- CRAN packages ---
cran_pkgs <- c("ape", "dplyr", "forcats", "scales", "ggplot2")
install.packages(setdiff(cran_pkgs, rownames(installed.packages())))

# --- Bioconductor core (needed by ggtree) ---
if (!requireNamespace("BiocManager", quietly = TRUE)) install.packages("BiocManager")
BiocManager::install(c("ggtree", "treeio", "tidytree"), update = FALSE, ask = FALSE)

# --- ggtreeExtra (Bioc; fall back to GitHub if needed) ---
if (!requireNamespace("ggtreeExtra", quietly = TRUE)) {
  try(BiocManager::install("ggtreeExtra", update = FALSE, ask = FALSE), silent = TRUE)
  if (!requireNamespace("ggtreeExtra", quietly = TRUE)) {
    if (!requireNamespace("remotes", quietly = TRUE)) install.packages("remotes")
    remotes::install_github("YuLab-SMU/ggtreeExtra")
  }}

library(ape); library(ggtree); library(ggtreeExtra); library(dplyr); library(forcats); library(scales); library (ggnewscale); library(rnaturalearth)

```

#######################
## Figure - MAP #######
#######################
```{r}
meta <- read.csv("study_info_map.csv")

world <- ne_countries(scale = "medium", returnclass = "sf")

# Prepare data
meta_count2 <- meta %>%
  mutate(
    lon_r = round(longitude, 2),
    lat_r = round(latitude, 2)
  ) %>%
  group_by(lon_r, lat_r, life_stage_tested, pathogen_type) %>%
  summarise(n_obs = n(), .groups = "drop") %>%
  # draw Bsal last so the (few) triangles sit on top of the Bd circles
  arrange(pathogen_type == "Bsal",
          factor(life_stage_tested,
                 levels = c("adult", "juvenile", "larva")))

# Plot
ggplot(data = world) +
  geom_sf(fill = "grey92", colour = "grey92") +
  geom_text(
    data = meta_count2,
    aes(x = lon_r, y = lat_r, label = n_obs),
    colour = "black",
    size = 2.4, fontface = "bold", vjust = 0.4, alpha = 0.8
  ) +
  geom_point(
    data = meta_count2,
    aes(x = lon_r, y = lat_r, shape = pathogen_type,
        fill = life_stage_tested, size = n_obs),
    colour = "black", stroke = 0.6
  ) +
  scale_size_area(
    name = "No. of Exp",
    max_size = 10,
    breaks = c(1, 3, 5), guide = "none"
  ) +
  scale_shape_manual(
    name = "Pathogen",
    values = c(Bd = 21, Bsal = 24),
    guide = "none"
  ) +
  scale_fill_manual(
    name = "Life Stage",
    values = scales::alpha(c(larva = "#0077B6", juvenile = "#E69F00", adult = "#D62828"), 0.5),
    guide = "none"
  ) +
  coord_sf(ylim = c(-90, 100), expand = FALSE) +
  theme_minimal() +
  theme(
    panel.background = element_rect(fill = "white"),
    panel.grid = element_line(colour = "gray90", linetype = "dotted"),
    legend.position = "none"
  ) +
  labs(x = NULL, y = NULL, title = NULL) +
  geom_rect(aes(xmin = -180, xmax = 180, ymin = -90, ymax = 100),
            fill = NA, colour = "black", linewidth = 0.3)

## output: 9 * 4 ## 5.61 * 1.9 inches (landscape)
```

#######################################################
## Main figure - MAP histogram count (per continent) ##
#######################################################
```{r}
histo <- read.csv("study_info_map.csv")

## Study and experiment count per continent
histo %>% group_by(continent) %>%summarise(n_study = n_distinct(study_id), .groups = "drop") 

# Step 0: create a count table
country_count <- histo %>% group_by(continent,treatment) %>%
  summarise(n_obs = n(), .groups = "drop") 

# Step 1: ensure all combinations exist (fill missing with 0)
df_complete <- country_count %>%
  tidyr::complete(continent, treatment, fill = list(n_obs = 0)) %>%
  mutate(
    treatment = factor(treatment,
      levels = c("Other", "Probiotics", "SynChem","NatChem","DeadFungi","LiveFungi")
    )
  ) %>%
  arrange(treatment)

# Split by continent
df_AU  <- df_complete %>% filter(continent == "AU")
df_EU  <- df_complete %>% filter(continent == "EU")
df_USA <- df_complete %>% filter(continent == "USA")

# Colour palette
treat_cols <- c(
  LiveFungi  = "#08306B",  # darkest blue
  DeadFungi  = "#08519C",
  NatChem    = "#2171B5",
  SynChem    = "#4292C6",
  Probiotics = "#6BAED6",
  Other      = "#C6DBEF"   # lightest blue
)
# ggplot style template
base_plot <- function(data, title) {
  ggplot(data, aes(y = treatment, x = n_obs, fill = treatment)) +   # swapped axes
    geom_col(colour = "black", width = 0.7) +
    scale_fill_manual(values = treat_cols) +scale_y_discrete(
      labels = c(
        LiveFungi  = "Live pathogen",
        DeadFungi  = "Dead pathogen",
        NatChem    = "Natural chemical",
        SynChem    = "Synthetic antiparasitics",
        Probiotics = "Probiotics",
        Other      = "Other"
      ))+
    coord_cartesian(xlim = c(0, 15)) +    # fixed x-axis range (was ylim)
    labs(title = title,
         x = "",
         y = "") +
    theme_minimal() +
    theme(
      axis.text.y = element_text(size = 12),          # clearer y-axis text
      axis.text.x = element_text(size = 12),
      legend.position = "none",
      panel.grid.minor = element_blank(),
      panel.grid.major = element_blank(),
      panel.background = element_rect(fill = NA, colour = "black", linewidth = 0.5),
      plot.background  = element_rect(fill = NA, colour = NA)
      )}

# Individual plots (output size: 5.18 * 1.92)
base_plot(df_AU,  "Oceania (n = 5 exp from 3 studies)")
base_plot(df_EU,  "Europe (n = 5 exp from 2 studies)")
base_plot(df_USA, "North America (n = 49 exp from 33 studies)")

## 5.61 * 1.9 inches (landscape)
```
```{r}
#########################################################
## Main figure - main phylogeny tree (with 23 species) ##
#########################################################
# Load the tree
tree <- read.tree("tree_for_imputation.tre")
amphib_tax <- read.csv("metadata.csv")

# Species set from my metadata
sp_data <- unique(trimws(gsub("\\s+", " ", amphib_tax$host_species)))
sp_data_treefmt <- gsub(" ", "_", sp_data)  # match typical tree tip style

# Check the mismatch between species in my metadata and species in existing tree
sp_tree_spaces <- gsub("_", " ", tree$tip.label)
setdiff(sp_data, sp_tree_spaces) # In data but not in tree

# Prune the tree: only keep species present in BOTH tree and my metadata
tips_to_keep <- intersect(tree$tip.label, sp_data_treefmt)
tree_pruned0 <- keep.tip(tree, tips_to_keep) 

# Force the tree to be binary
tree_pruned <- multi2di(tree_pruned0)
is.binary (tree_pruned) 

# Create a column in metadata that matches the tree’s tip-label format
amphib_tax$species <- gsub(" ", "_", trimws(amphib_tax$host_species))

# tidy ordering (apply in place so it actually takes effect)
tree_pruned <- ape::ladderize(tree_pruned)

# Replace underscores in species names
tree_pruned$tip.label <- gsub("_", " ", tree_pruned$tip.label)

##########################
#### A Basic tree (p) ####
##########################
p <- ggtree( tree_pruned, layout = "rectangular", branch.length = "none") + geom_tree(linewidth = 0.4) + theme( plot.margin = margin(5, 5, 5, 5), legend.position = "none") + xlim (0, 15)

p 

#######################################
### Tree (p) + species name = p2 ######
#######################################
p2 <- p +
  geom_tiplab(aes(label = label), size = 7, offset = 0.2, align = FALSE, fontface = "italic", family = "sans") +
  theme(plot.margin = margin(50, 50, 50, 50))

p2 

p2 + theme(aspect.ratio = 3)   # height:width = 2:1; raise for more vertical spread

# Export = 25 * 35 inches
```

###################################################################
## Effect size figures - Overall effect (Extended Data Figure 2) ##
###################################################################
```{r}
# Note that all models using here are all from the R script 'Meta_analyis_code'. You should ran that script before making figures here.

###################
## (1) mortality ##
###################
Mor_all_plot <- orchard_plot(
  M3_mor,
  group = "study_exp_id",
  xlab  = "Log risk ratio",
  angle = 0,
  branch.size = 1.3,
  trunk.size = 0.7,
  twig.size = 0, cb = FALSE, colour = FALSE, fill = TRUE, alpha = 0.7)

# grey points, no jitter, larger min size
idx <- which(vapply(Mor_all_plot$layers, \(x) inherits(x$geom, "GeomPoint"), logical(1)))[1]
if (!is.na(idx)) Mor_all_plot$layers[[idx]]$aes_params[c("colour","fill","alpha")] <- list("grey80","grey60",0.2)

# move "k = ..." text down to same line as intercept
k_idx <- which(vapply(Mor_all_plot$layers, \(x) inherits(x$geom, "GeomText"), logical(1)))[1]
if (!is.na(k_idx))
  Mor_all_plot$layers[[k_idx]]$position <- position_nudge(x = -0.3, y = 0.5)  # adjust as needed

Mor_all_plot + scale_x_discrete(labels = NULL) + scale_size_continuous(range = c(4, 8)) + guides(fill = "none", colour = "none", size = "none") + theme(legend.position = c(0.57, 0.82),panel.grid.minor = element_blank(), panel.grid.major = element_blank()) + geom_hline(yintercept = 0, linetype = "dashed", color = "grey50")

## output: 5.67 * 1.84 inches (landscape)

####################
## (2) prevalence ##
####################

Preval_all_plot <- orchard_plot(
  M3_preval,
  group = "study_exp_id",
  xlab  = "Log risk ratio",
  angle = 0,
  branch.size = 1.3,
  trunk.size = 0.7,
  twig.size = 0, cb = FALSE, colour = FALSE, fill = TRUE, alpha = 0.7)

Preval_all_plot

# grey points, no jitter, larger min size
idx <- which(vapply(Preval_all_plot$layers, \(x) inherits(x$geom, "GeomPoint"), logical(1)))[1]
if (!is.na(idx)) Preval_all_plot$layers[[idx]]$aes_params[c("colour","fill","alpha")] <- list("grey80","grey60",0.2)

# move "k = ..." text down to same line as intercept
k_idx <- which(vapply(Preval_all_plot$layers, \(x) inherits(x$geom, "GeomText"), logical(1)))[1]
if (!is.na(k_idx))
  Preval_all_plot$layers[[k_idx]]$position <- position_nudge(x = -0.3, y = 0.08)  # adjust as needed

Preval_all_plot + scale_x_discrete(labels = NULL) + scale_size_continuous(range = c(4, 8)) + guides(fill = "none", colour = "none", size = "none") + theme(legend.position = c(0.57, 0.82),panel.grid.minor = element_blank(), panel.grid.major = element_blank()) + geom_hline(yintercept = 0, linetype = "dashed", color = "grey50")

###################
## (3) intensity ##
###################

LOAD_all_plot <- orchard_plot(
  M3_INC,
  group = "study_exp_id",
  xlab  = "Log response ratio",
  angle = 0,
  branch.size = 1.3,
  trunk.size = 0.7,
  twig.size = 0, cb = FALSE, colour = FALSE, fill = TRUE, alpha = 0.7)

LOAD_all_plot

# grey points, no jitter, larger min size
idx <- which(vapply(LOAD_all_plot$layers, \(x) inherits(x$geom, "GeomPoint"), logical(1)))[1]
if (!is.na(idx)) LOAD_all_plot$layers[[idx]]$aes_params[c("colour","fill","alpha")] <- list("grey80","grey60",0.2)

# move "k = ..." text down to same line as intercept
k_idx <- which(vapply(LOAD_all_plot$layers, \(x) inherits(x$geom, "GeomText"), logical(1)))[1]
if (!is.na(k_idx))
  LOAD_all_plot$layers[[k_idx]]$position <- position_nudge(x = -0.3, y = 0.8)  # adjust as needed

LOAD_all_plot + scale_x_discrete(labels = NULL) + scale_size_continuous(range = c(4, 8)) + guides(fill = "none", colour = "none", size = "none") + theme(legend.position = c(0.57, 0.82),panel.grid.minor = element_blank(), panel.grid.major = element_blank()) + geom_hline(yintercept = 0, linetype = "dashed", color = "grey50")
```

##########################################################################
## Effect size figures - Moderating effect of treatment type (Figure 3) ##
##########################################################################
```{r}
# Note that all models using here are all from the R script 'Meta_analyis_code'. You should ran that script before making figures here.

# Set colours for treatment types

treat_cols <- c(
  "Live_pathogen"            = "#E69F00",
  "Dead_pathogen"            = "#56B4E9",
  "Natural_chem"             = "#009E73",
  "Synthetic_antiparastics"  = "#F0E442",
  "Probiotic"                = "#0072B2",
  "Other"                    = "#D55E00"
)

###################
## (1) Mortality ##
###################

Mor_T_plot <- orchard_plot(
  M_treat_dead1,
  mod = "treatment",
  group = "study_exp_id",
  xlab  = "",
  angle = 0,
  twig.size = 0, 
  branch.size = 1.3,
  trunk.size = 0.7,
  cb = FALSE, colour = FALSE, fill = TRUE, alpha = 0.1)
Mor_T_plot
# move "k = ..." text down to same line as intercept
k_idx <- which(vapply(Mor_T_plot$layers, \(x) inherits(x$geom, "GeomText"), logical(1)))[1]
if (!is.na(k_idx))
  Mor_T_plot$layers[[k_idx]]$position <- position_nudge(x = -0.3, y = 0.9)  # adjust as needed

Mor_T_plot + geom_hline(yintercept = 0, linetype = "dashed", color = "grey50") +
  scale_fill_manual(values = treat_cols, name = "Treatment", drop = FALSE) +
  scale_colour_manual(values = treat_cols, name = "Treatment", drop = FALSE) +
  scale_size_continuous(range = c(4.5, 8)) +
  guides(colour = "none", fill ="none") +                           # keep only one legend if you like
  theme(legend.position = c(-1, 0.20),panel.grid.minor = element_blank(), panel.grid.major = element_blank(),) + scale_x_discrete(labels = NULL)

## 5 * 3 inches

##############################
## (2) Infection prevalence ##
##############################

prev_T_plot <- orchard_plot(
  M_treat_infected,
  mod = "treatment",
  group = "study_exp_id",
  xlab  = "",
  angle = 0,
  twig.size = 0, 
  branch.size = 1.3,
  trunk.size = 0.7,
  cb = FALSE, colour = FALSE, fill = TRUE, alpha = 0.1)

prev_T_plot

# move "k = ..." text down to same line as intercept
k_idx <- which(vapply(prev_T_plot$layers, \(x) inherits(x$geom, "GeomText"), logical(1)))[1]
if (!is.na(k_idx))
  prev_T_plot$layers[[k_idx]]$position <- position_nudge(x = -0.3, y = 0.9)  # adjust as needed

prev_T_plot + geom_hline(yintercept = 0, linetype = "dashed", color = "grey50") +
  scale_fill_manual(values = treat_cols, name = "Treatment", drop = FALSE) +
  scale_colour_manual(values = treat_cols, name = "Treatment", drop = FALSE) +
  scale_size_continuous(range = c(4.5, 8)) +
  guides(colour = "none", size = "none", fill ="none") +                           # keep only one legend if you like
  theme(legend.position = c(0.57, 0.82),panel.grid.minor = element_blank(), panel.grid.major = element_blank(),) + scale_x_discrete(labels = NULL)

## 6.01 * 3 inches

#############################
## (3) Infection intensity ##
#############################

Load_T_plot <- orchard_plot(
  M_T_INC,
  mod = "treatment",
  group = "study_exp_id",
  xlab  = "",
  angle = 0,
  twig.size = 0, 
  branch.size = 1.3,
  trunk.size = 0.7,
  cb = FALSE, colour = FALSE, fill = TRUE, alpha = 0.1)

Load_T_plot

# move "k = ..." text down to same line as intercept
k_idx <- which(vapply(Load_T_plot$layers, \(x) inherits(x$geom, "GeomText"), logical(1)))[1]
if (!is.na(k_idx))
  Load_T_plot$layers[[k_idx]]$position <- position_nudge(x = -0.3, y = 0.9)  # adjust as needed

Load_T_plot + geom_hline(yintercept = 0, linetype = "dashed", color = "grey50") +
  scale_fill_manual(values = treat_cols, name = "Treatment", drop = FALSE) +
  scale_colour_manual(values = treat_cols, name = "Treatment", drop = FALSE) +
  scale_size_continuous(range = c(4.5, 8)) +
  guides(colour = "none", size = "none", fill ="none") +                           # keep only one legend if you like
  theme(legend.position = c(0.57, 0.82),panel.grid.minor = element_blank(), panel.grid.major = element_blank(),) + scale_x_discrete(labels = NULL)

## 6.01 * 3 inches
```

#######################################################################
## Effect size figures - Moderating effect of life stage (Figure 4) ###
#######################################################################
```{r}
# Note that all models using here are all from the R script 'Meta_analyis_code'. You should ran that script before making figures here.


# Set Colours for life stage
stage_cols <- c(
  "Adult"       = "#D62828",
  "Juvenile"    = "#FFB703",
  "Larva"        = "#0077B6")

###################
## (1) mortality ##
###################

Mor_S_plot <- orchard_plot(
  M_stage_dead,
  mod = "life_stage_tested",
  group = "study_exp_id",
  xlab  = "Log risk ratio",
  angle = 0,
  twig.size = 0, 
  branch.size = 1.3,
  trunk.size = 0.7,
  cb = FALSE, colour = FALSE, fill = TRUE, alpha = 0.1)
Mor_S_plot
# move "k = ..." text down to same line as intercept
k_idx <- which(vapply(Mor_S_plot$layers, \(x) inherits(x$geom, "GeomText"), logical(1)))[1]
if (!is.na(k_idx))
  Mor_S_plot$layers[[k_idx]]$position <- position_nudge(x = -0.3, y = 5.6)  # adjust as needed

Mor_S_plot + geom_hline(yintercept = 0, linetype = "dashed", color = "grey50") +
  scale_fill_manual(values = stage_cols, name = "life_stage_tested", drop = FALSE) +
  scale_colour_manual(values = stage_cols, name = "life_stage_tested", drop = FALSE) +
  scale_size_continuous(range = c(2, 8)) +
  guides(colour = "none", fill ="none", size = "none") +                           # keep only one legend if you like
  theme(legend.position = c(0.5, 0.20),panel.grid.minor = element_blank(), panel.grid.major = element_blank(),) + scale_x_discrete(labels = NULL)+ylim (-3.5, 8)

## 3.2 * 3 inches (protrait)

##############################
## (2) Infection prevalence ##
##############################

prev_S_plot <- orchard_plot(
  M_stage_infected,
  mod = "life_stage_tested",
  group = "study_exp_id",
  xlab  = "Log risk ratio",
  angle = 0,
  twig.size = 0, 
  branch.size = 1.3,
  trunk.size = 0.7,
  cb = FALSE, colour = FALSE, fill = TRUE, alpha = 0.1)

prev_S_plot
# move "k = ..." text down to same line as intercept
k_idx <- which(vapply(prev_S_plot$layers, \(x) inherits(x$geom, "GeomText"), logical(1)))[1]
if (!is.na(k_idx))
  prev_S_plot$layers[[k_idx]]$position <- position_nudge(x = -0.3, y = 0.7)  # adjust as needed

prev_S_plot + geom_hline(yintercept = 0, linetype = "dashed", color = "grey50") +
  scale_fill_manual(values = stage_cols, name = "life_stage_tested", drop = FALSE) +
  scale_colour_manual(values = stage_cols, name = "life_stage_tested", drop = FALSE) +
  scale_size_continuous(range = c(2, 8)) +
  guides(colour = "none", fill ="none", size = "none") +                           # keep only one legend if you like
  theme(legend.position = c(0.5, 0.20),panel.grid.minor = element_blank(), panel.grid.major = element_blank(),) + scale_x_discrete(labels = NULL) + ylim (-3.1,3.1)

## 3.2 * 3 inches (protrait)

#############################
## (3) Infection intensity ##
#############################

Load_S_plot <- orchard_plot(
  M_LS_INC,
  mod = "life_stage_tested",
  group = "study_exp_id",
  xlab  = "Log response ratio",
  angle = 0,
  twig.size = 0, 
  branch.size = 1.3,
  trunk.size = 0.7,
  cb = FALSE, colour = FALSE, fill = TRUE, alpha = 0.1)

Load_S_plot
# move "k = ..." text down to same line as intercept
k_idx <- which(vapply(Load_S_plot$layers, \(x) inherits(x$geom, "GeomText"), logical(1)))[1]
if (!is.na(k_idx))
  Load_S_plot$layers[[k_idx]]$position <- position_nudge(x = -0.3, y = 2) # adjust as needed

Load_S_plot + geom_hline(yintercept = 0, linetype = "dashed", color = "grey50") +
  scale_fill_manual(values = stage_cols, name = "life_stage_tested", drop = FALSE) +
  scale_colour_manual(values = stage_cols, name = "life_stage_tested", drop = FALSE) +
  scale_size_continuous(range = c(2, 8)) +
  guides(colour = "none", fill ="none", size = "none") +                           # keep only one legend if you like
  theme(legend.position = c(0.5, 0.20),panel.grid.minor = element_blank(), panel.grid.major = element_blank(),) + scale_x_discrete(labels = NULL) + ylim(-8.1,8.1)

## 3.2 * 3 inches (protrait)

```
#######################################################################
## Effect size figures - Moderating effect of host origin (Figure 4) ##
#######################################################################
```{r}
# Note that all models using here are all from the R script 'Meta_analyis_code'. You should rUn that script before making figures here.



# Set colours for host origins
origin_cols <- c(
  "Captive"     = "#6A4C93",
  "Wild"  = "#2A9D8F")

###################
## (1) mortality ##
###################

Mor_O_plot <- orchard_plot(
  M_host_dead1,
  mod = "host_origin",
  group = "study_exp_id",
  xlab  = "Log risk ratio",
  angle = 0,
  twig.size = 0, 
  branch.size = 1.3,
  trunk.size = 0.7,
  cb = FALSE, colour = FALSE, fill = TRUE, alpha = 0.1)
Mor_O_plot

# move "k = ..." text down to same line as intercept
k_idx <- which(vapply(Mor_O_plot$layers, \(x) inherits(x$geom, "GeomText"), logical(1)))[1]
if (!is.na(k_idx))
  Mor_O_plot$layers[[k_idx]]$position <- position_nudge(x = -0.3, y = 3)  # adjust as needed

Mor_O_plot + geom_hline(yintercept = 0, linetype = "dashed", color = "grey50") +
  scale_fill_manual(values = origin_cols, name = "life_stage_tested", drop = FALSE) +
  scale_colour_manual(values = origin_cols, name = "life_stage_tested", drop = FALSE) +
  scale_size_continuous(range = c(2, 6)) +
  guides(colour = "none", fill ="none", size = "none") +                           # keep only one legend if you like
  theme(legend.position = c(0.5, 0.20),panel.grid.minor = element_blank(), panel.grid.major = element_blank(),) + scale_x_discrete(labels = NULL) + ylim(-3.5, 5.4)

## 3.2 * 2 inches (landscape)

####################
## (2) prevalence ##
####################

Preval_O_plot <- orchard_plot(
  M_host_infected,
  mod = "host_origin",
  group = "study_exp_id",
  xlab  = "Log risk ratio",
  angle = 0,
  twig.size = 0, 
  branch.size = 1.3,
  trunk.size = 0.7,
  cb = FALSE, colour = FALSE, fill = TRUE, alpha = 0.1)
Preval_O_plot
# move "k = ..." text down to same line as intercept
k_idx <- which(vapply(Preval_O_plot$layers, \(x) inherits(x$geom, "GeomText"), logical(1)))[1]
if (!is.na(k_idx))
  Preval_O_plot$layers[[k_idx]]$position <- position_nudge(x = -0.3, y = 0.5)  # adjust as needed

Preval_O_plot + geom_hline(yintercept = 0, linetype = "dashed", color = "grey50") +
  scale_fill_manual(values = origin_cols, name = "life_stage_tested", drop = FALSE) +
  scale_colour_manual(values = origin_cols, name = "life_stage_tested", drop = FALSE) +
  scale_size_continuous(range = c(2, 6)) +
  guides(colour = "none", fill ="none", size = "none") +                           # keep only one legend if you like
  theme(legend.position = c(0.5, 0.20),panel.grid.minor = element_blank(), panel.grid.major = element_blank(),) + scale_x_discrete(labels = NULL)

## 3.2 * 2 inches (landscape)

###################
## (3) intensity ##
###################

Load_O_plot <- orchard_plot(
  M_origin_juv_INC,
  mod = "host_origin",
  group = "study_exp_id",
  xlab  = "Log response ratio",
  angle = 0,
  twig.size = 0, 
  branch.size = 1.3,
  trunk.size = 0.7,
  cb = FALSE, colour = FALSE, fill = TRUE, alpha = 0.1)
Load_O_plot
# move "k = ..." text down to same line as intercept
k_idx <- which(vapply(Load_O_plot$layers, \(x) inherits(x$geom, "GeomText"), logical(1)))[1]
if (!is.na(k_idx))
  Load_O_plot$layers[[k_idx]]$position <- position_nudge(x = -0.3, y = 4.4)  # adjust as needed

Load_O_plot + geom_hline(yintercept = 0, linetype = "dashed", color = "grey50") +
  scale_fill_manual(values = origin_cols, name = "life_stage_tested", drop = FALSE) +
  scale_colour_manual(values = origin_cols, name = "life_stage_tested", drop = FALSE) +
  scale_size_continuous(range = c(2, 6)) +
  guides(colour = "none", fill ="none", size = "none") +                           # keep only one legend if you like
  theme(legend.position = c(0.5, 0.20),panel.grid.minor = element_blank(), panel.grid.major = element_blank(),) + scale_x_discrete(labels = NULL) + ylim (-8.1, 8.1)

## 3.2 * 2 inches (landscape)
```
#########################################################################
## Effect size figures - Moderating effect of host taxonomy (Figure 4) ##
#########################################################################

# Note that all models using here are all from the R script 'Meta_analyis_code'. You should run that script before making figures here.

```{r}
# Set colours for host taxonomy (we only present three families with sufficient sample sizes)

family_cols <- c(
  "Bufonidae" = "#A7BE76",
  "Hylidae"   = "#E49E88",
  "Ranidae"   = "#CCA280")

###################
## (1) mortality ##
###################

Mor_t_plot <- orchard_plot(
  M_taxa_dead1,
  mod = "family",
  group = "study_exp_id",
  xlab  = "Log risk ratio",
  angle = 0,
  twig.size = 0, 
  branch.size = 1.3,
  trunk.size = 0.7,
  cb = FALSE, colour = FALSE, fill = TRUE, alpha = 0.1)
Mor_t_plot

# move "k = ..." text down to same line as intercept
k_idx <- which(vapply(Mor_t_plot$layers, \(x) inherits(x$geom, "GeomText"), logical(1)))[1]
if (!is.na(k_idx))
  Mor_t_plot$layers[[k_idx]]$position <- position_nudge(x = -0.3, y = 3)  # adjust as needed

Mor_t_plot + geom_hline(yintercept = 0, linetype = "dashed", color = "grey50") +
  scale_fill_manual(values = family_cols, name = "life_stage_tested", drop = FALSE) +
  scale_colour_manual(values = family_cols, name = "life_stage_tested", drop = FALSE) +
  scale_size_continuous(range = c(2, 6)) +
  guides(colour = "none", fill ="none", size = "none") +                           # keep only one legend if you like
  theme(legend.position = c(0.5, 0.20),panel.grid.minor = element_blank(), panel.grid.major = element_blank(),) + scale_x_discrete(labels = NULL) + ylim(-3.5,5.4)
  
### 3.2 *7 inches (protrait) >> 7 because each family (n = 7) took 1 space 

####################
## (2) prevalence ##
####################

prev_t_plot <- orchard_plot(
  M_taxa_infected,
  mod = "family",
  group = "study_exp_id",
  xlab  = "Log risk ratio",
  angle = 0,
  twig.size = 0, 
  branch.size = 1.3,
  trunk.size = 0.7,
  cb = FALSE, colour = FALSE, fill = TRUE, alpha = 0.1)
prev_t_plot

# move "k = ..." text down to same line as intercept
k_idx <- which(vapply(prev_t_plot$layers, \(x) inherits(x$geom, "GeomText"), logical(1)))[1]
if (!is.na(k_idx))
  prev_t_plot$layers[[k_idx]]$position <- position_nudge(x = -0.3, y = 0.7)  # adjust as needed

prev_t_plot + geom_hline(yintercept = 0, linetype = "dashed", color = "grey50") +
  scale_fill_manual(values = family_cols, name = "life_stage_tested", drop = FALSE) +
  scale_colour_manual(values = family_cols, name = "life_stage_tested", drop = FALSE) +
  scale_size_continuous(range = c(2, 6)) +
  guides(colour = "none", fill ="none", size = "none") +                           # keep only one legend if you like
  theme(legend.position = c(0.5, 0.20),panel.grid.minor = element_blank(), panel.grid.major = element_blank(),) + scale_x_discrete(labels = NULL) + ylim (-3.1,3.1)

### 3.2 *7 inches (protrait) >> 7 because each family (n = 7) took 1 space 

###################
## (3) intensity ##
###################

Load_t_plot <- orchard_plot(
  M_taxa_INC,
  mod = "family",
  group = "study_exp_id",
  xlab  = "Log response ratio",
  angle = 0,
  twig.size = 0, 
  branch.size = 1.3,
  trunk.size = 0.7,
  cb = FALSE, colour = FALSE, fill = TRUE, alpha = 0.1)
Load_t_plot

# move "k = ..." text down to same line as intercept
k_idx <- which(vapply(Load_t_plot$layers, \(x) inherits(x$geom, "GeomText"), logical(1)))[1]
if (!is.na(k_idx))
  Load_t_plot$layers[[k_idx]]$position <- position_nudge(x = -0.3, y = 6) # adjust as needed

Load_t_plot + geom_hline(yintercept = 0, linetype = "dashed", color = "grey50") +
  scale_fill_manual(values = family_cols, name = "life_stage_tested", drop = FALSE) +
  scale_colour_manual(values = family_cols, name = "life_stage_tested", drop = FALSE) +
  scale_size_continuous(range = c(2, 6)) +
  guides(colour = "none", fill ="none", size = "none") +                           # keep only one legend if you like
  theme(legend.position = c(0.5, 0.20),panel.grid.minor = element_blank(), panel.grid.major = element_blank(),) + scale_x_discrete(labels = NULL) + ylim(-12,12)


### 3.2 *9 inches (protrait) >> 7 because each family (n = 9) took 1 space 

```

##########################################################
## Absolute levels of prevalence (supplementary figure) ##
##########################################################
```{r}
### Absolute level for prevalence per group per treatment

## read.table
metadata <- read.csv("metadata.csv")
prevalence <- metadata %>% filter(trait == "prevalence")

df_long <- prevalence %>%
  mutate(row_id = row_number()) %>%   # if you don't already have an ID
  pivot_longer(
    cols = c(mean_C_trait_initial, mean_T_trait_initial),
    names_to = "group",
    values_to = "prevalence"
  ) %>%
  mutate(group = recode(group,
                        "mean_C_trait_initial" = "Untreated",
                        "mean_T_trait_initial" = "Treated"))

df_long$treatment <- factor(
  df_long$treatment,
  levels = c(
    "live_pathogen",
    "natural_chem",
    "synthetic_antiparastics",
    "dead_pathogen",
    "probiotic",
    "other"
  )
)

df_summary <- df_long %>%
  group_by(treatment, group) %>%
  summarise(
    mean_prev = mean(prevalence, na.rm = TRUE),
    se_prev = sd(prevalence, na.rm = TRUE) / sqrt(n()),
    .groups = "drop"
  )



df_long$group <- factor(df_long$group,
                        levels = c("Untreated", "Treated"))

ggplot(df_long,
       aes(x = group, y = prevalence,
           group = row_id)) +
  
  # raw paired lines + points
  geom_line(alpha = 0.3, colour = "grey40") +
  geom_point(size = 2, alpha = 0.7, position = position_jitter(width = 0.03, height = 0), colour = "black") +
  
  # --- NEW: treatment-specific mean + SE ---
  geom_point(data = df_summary,
             aes(x = group, y = mean_prev),
             size = 4,
             colour = "red",
             inherit.aes = FALSE) +
  
  geom_errorbar(data = df_summary,
                aes(x = group,
                    ymin = mean_prev - se_prev,
                    ymax = mean_prev + se_prev),
                width = 0.15,
                colour = "red",
                inherit.aes = FALSE) +
  # -----------------------------------------

facet_wrap(~ treatment, nrow = 1,
           labeller = as_labeller(c(
             live_pathogen = "Live pathogen",
             dead_pathogen = "Dead pathogen",
             natural_chem = "Natural chemical",
             synthetic_antiparastics = "Syn antiparasitic",
             probiotic = "Probiotic",
             other = "Other"
           ))) +
  labs(
    x = "",
    y = "Prevalence (%)"
  ) +
  theme_bw() +
  theme(
    strip.text = element_text(size = 12, face = "bold"),
    panel.grid.minor = element_blank(),
    panel.grid.major = element_line(colour = "aliceblue", linewidth = 0.02)
  )
```


#########################################################
## Absolute levels of mortality (supplementary figure) ##
#########################################################

```{r}
## read.table
metadata <- read.csv("metadata.csv")
mortality <- metadata %>% filter(trait == "mortality")

df_long <- mortality %>%
  mutate(row_id = row_number()) %>%   # if you don't already have an ID
  pivot_longer(
    cols = c(mean_C_trait_initial, mean_T_trait_initial),
    names_to = "group",
    values_to = "mortality"
  ) %>%
  mutate(group = recode(group,
                        "mean_C_trait_initial" = "Untreated",
                        "mean_T_trait_initial" = "Treated"))

df_long$treatment <- factor(
  df_long$treatment,
  levels = c(
    "live_pathogen",
    "natural_chem",
    "synthetic_antiparastics",
    "dead_pathogen",
    "probiotic",
    "other"
  )
)

df_summary <- df_long %>%
  group_by(treatment, group) %>%
  summarise(
    mean_mor = mean(mortality, na.rm = TRUE),
    se_mor = sd(mortality, na.rm = TRUE) / sqrt(n()),
    .groups = "drop"
  )



df_long$group <- factor(df_long$group,
                        levels = c("Untreated", "Treated"))

ggplot(df_long,
       aes(x = group, y = mortality,
           group = row_id)) +
  
  # raw paired lines + points
  geom_line(alpha = 0.3, colour = "grey40") +
  geom_point(size = 2, alpha = 0.7, position = position_jitter(width = 0.03, height = 0), colour = "black") +
  
  # --- NEW: treatment-specific mean + SE ---
  geom_point(data = df_summary,
             aes(x = group, y = mean_mor),
             size = 4,
             colour = "red",
             inherit.aes = FALSE) +
  
  geom_errorbar(data = df_summary,
                aes(x = group,
                    ymin = mean_mor - se_mor,
                    ymax = mean_mor + se_mor),
                width = 0.15,
                colour = "red",
                inherit.aes = FALSE) +
  # -----------------------------------------

facet_wrap(~ treatment, nrow = 1,
           labeller = as_labeller(c(
             live_pathogen = "Live pathogen",
             dead_pathogen = "Dead pathogen",
             natural_chem = "Natural chemical",
             synthetic_antiparastics = "Syn antiparasitic",
             probiotic = "Probiotic",
             other = "Other"
           ))) +
  labs(
    x = "",
    y = "Mortality rate (%)"
  ) +
  theme_bw() +
  theme(
    strip.text = element_text(size = 12, face = "bold"),
    panel.grid.minor = element_blank(),
    panel.grid.major = element_line(colour = "aliceblue", linewidth = 0.02)
  )
```
